---
title: "Sudan_catch_analysis"
author: "Erik Olsen"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output:
  html_document: default
  pdf_document: default
---
# Analysis of Sudan survey data 2012 - 2013

This R-markdown document contains all r-code used to carry out the analysis for the paper on fish catch rates and biodiversity along the Red Sea coast of Sudan. The analysis is based on three surveys: November 2012, May 2013 and November 2013. 

Here the analysis is structured around the 7 management regions of the Sudanese Red Sea coast as follows:

- Present the region: Bathymetric map
- Sampling locations
- Table of effort (diffent gear types per management areas.)
- Table of species / families caught
- Depth distribution of traps by survey and area
- Catch CPUE by area and survey
- Catch CPUE distribution by gear, by species groups (should perhaps do the same for trophic groups)
- PCA analysis of species and trophic groups in catches
- CPUE by trophic group pr area and survey
- CPUE by depth and tropich group and survey
- Statistical analysis
  - test of normality shows that we need to use non-parametric and zero-adjusted approaches. 
  - CPUE as a function of area, depth, survey, gear (Zero-adjusted model)
- Biodiversity analysis
  - Species accumulation curves
  - Functional diversity (Rao´s Q)
  
All code and data are stored on GitHub

## Libraries, data etc. 
### Libraries, dependencies and functions
```{r libraries and dependencies, echo=FALSE, warning=FALSE, message=FALSE}
library(reshape2)
library(maps)
library(mapdata)
library(ggplot2)
library(RColorBrewer)
library(sp)
library(rgdal)
library(rgeos)
# plyr mus be loaded befor dplyr for both to work
library(plyr) 
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(maptools)
library(marmap)
library(classInt)
library(colorspace)
library("formattable")
library("PMCMR")
library(FD)
library("gamlss")
library(patchwork)
```


```{r functions, echo=FALSE, warning=FALSE}

capwords <- function(s, strict = FALSE) {
  cap <- function(s) paste(toupper(substring(s, 1, 1)),
                           {s <- substring(s, 2); if(strict) tolower(s) else s},
                           sep = "", collapse = " " )
  sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
}
```


### Load and manipulate data
Reading catch, station and traits data. 

Neither station data, nor catch data has a complete depth record, but by combining depth para from each we can get a complete depth parameter for all stations. 
```{r load data, echo=FALSE, warning=FALSE}
#' Catch data
catch<-read.csv2("catch.csv")
catch <- catch[,c(1:15, 17:24)]
catch$survey<-as.factor(catch$survey)
catch$CPUEw<-catch$weight/catch$Fhrs
catch$CPUEn<-catch$number/catch$Fhrs

#' Station
station <- read.csv2("stations.csv")

#' add depth to catch from station
catch$d1<- station$geardepthstart[match(catch$ss, station$ss)]  
catch$depth <- apply(X=cbind(catch$bdep, catch$d1), MARGIN=1, FUN=max, na.rm=TRUE)

#' Add Month. Year survey column
catch$survey_m<-as.character(catch$survey)
catch$survey_m<-replace(catch$survey_m, grep("2012901", catch$survey_m), "Survey 1: Nov. 2012")
catch$survey_m<-replace(catch$survey_m, grep("2013002", catch$survey_m), "Survey 2: May 2013")
catch$survey_m<-replace(catch$survey_m, grep("2013005", catch$survey_m), "Survey 3: Nov. 2013")

#'Traits
traits <- read.csv2("traits_noNa.csv")
```


### Sudan map and management areas
Loads the map data for Sudan, loads the management areas from shape-file etc. 
```{r map data, echo=FALSE, warning=FALSE, message=FALSE}
#' SUDAN MAP DATA
world<- map_data("worldHires", c("Sudan", "Ethiopia")) 
world$survey_m<-c("Survey 1: Nov. 2012")
world2<-world
world2$survey_m<-c("Survey 2: May 2013")
world3<-world
world3$survey_m<-c("Survey 3: Nov. 2013")
world<-rbind(world, world2, world3)

#'  SPATIAL POLYGONS FOR MANAGEMENT AREAS
#setwd("/Users/eriko/Documents/GitHub/Sudan2019/Sudan-master")
setwd("/Users/eriko/GitHub/Sudan2019/Sudan-master")
setwd("./sudan_management_areas")
ManageAreas<-readOGR(dsn = ".", "sudan_regions")
setwd("..")

#' setting correct projection LAT LON
ManageAreas <- spTransform(ManageAreas, CRS("+proj=longlat +ellps=GRS80"))

#' Create  area map with numbers for each region
ManageAreas.f<-fortify(ManageAreas, region="id") #creates X - Y points of the polygons
cnames <- aggregate(cbind(long, lat) ~ id, data=ManageAreas.f, FUN=function(x)mean(range(x)))
cnames$id<-c(1:7)

#' position of Port Sudan
ps<-rbind(data.frame(x=37.21709967,y=19.600512, group=1, survey_m=c("Survey 1: Nov. 2012")),data.frame(x=37.21709967,y=19.600512, group=1, survey_m=c("Survey 2: May 2013") ),data.frame(x=37.21709967,y=19.600512, group=1, survey_m=c("Survey 3: Nov. 2013") ))
```

### Allocating catch positions to management areas
Adds 'Area' to each line in the catch table. 
```{r catch pos to mngmt areas, echo=FALSE, warning=FALSE, message=FALSE}
#' Create spatial.points data base from catch
catch.points<- SpatialPoints(catch[13:14])
proj4string(catch.points) <- proj4string(ManageAreas)
catch.areas<-over(catch.points, ManageAreas)

#' fix area number to area name
alist<-data.frame(name=levels(as.factor(catch.areas$name)), id=c(7,3,2,1,4,5,6))
catch <- cbind(catch, catch.areas[2])
catch$id <- alist$id[match(catch$name, alist$name)]  

#' allocate management areas to catches outside of polygons (choose nearest polygon)
naarea <- catch[is.na(catch$id),]
naarea$id <- naarea$Area
naarea$name[1] <- c("Arakiai")
naarea$name[2:8] <- c("Suakin Archipelago")
naarea$name[9:16] <- c("Donganab")
catch<- rbind(subset(catch, id>=1), naarea) 

#' Change family names to Upper & Lower case
catch$FamGroup2 <- paste(toupper(substr(catch$FamGroup, 1, 1)), tolower(substr(catch$FamGroup, 2, 20)), sep="")
catch$FamGroup2 <- as.factor(catch$FamGroup2)
```




### Modifying dataset and creating one dataset for CPUE analysis and another for traits
Of the catch data, 12 fish registrations lack weight, (i.e. this was forgotten entered into the database during the survey). These registrations, would if included lead to 12 more registrations of CPUE = 0. 
```{r nocatch dataset, echo=FALSE, warning=FALSE}
#' Keep a copy of original 'catch'
catch.unmod <- catch

#' remove Hand Line stations with too short time <1 hr, and too long time >8 hrs
# hist((subset(catch, gear=="HL" & Fhrs>0))$Fhrs, breaks =20) # get an idea of the spread in fishing time
catch <- rbind(subset(catch, gear!="HL"), subset(catch, gear=="HL" & Fhrs>1 & Fhrs<8))

#' remove 14 records where weight of fish has not been recorded
names <- row.names(subset(catch, weight==0 & number>0))
catch.cpue <- subset(catch, !(rownames(catch) %in% names))


#' selecting only stations with catch & removes stations shorter than 1 hour - two handline stations with fishing time recorded as 5min, that we know is wrong, but we don't have the correct time. 
catch1<-subset(catch, Sci_name!="No catch" & Sci_name!="NO CATCH") 

```

### Adding traits to 'catch' data set (without 0-catch stations)
Some more data wrangling that adds the traits from the traits table to the catch data, using only the station with catches (there are no traits for 'NOCATCH' species). 

Also, only select traps, gillnets and handlines as these were the only gear with sufficient numbers and consistent use to be analyzed. 

Lastly adds number of gear deployed at each station to each line. 
```{r add traits,  echo=FALSE, warning=FALSE, message=FALSE}
catch.traits <- catch1
catch.traits$MaxLength<- traits$MaxLength[match(catch1$Sci_name, traits$x1)]  
catch.traits$TrophicLevel<- traits$Trophic.Level[match(catch1$Sci_name, traits$x1)]  
catch.traits$TrophicGroup<- traits$Trophic.group[match(catch1$Sci_name, traits$x1)] 
catch.traits$WaterCol<- traits$Water.column[match(catch1$Sci_name, traits$x1)] 
catch.traits$DielActivity<- traits$Diel.Activity[match(catch1$Sci_name, traits$x1)] 
catch.traits$Habitat<- traits$Habitat[match(catch1$Sci_name, traits$x1)] 
catch.traits$Gregariousness<- traits$Gregariousness[match(catch1$Sci_name, traits$x1)] 
catch.traits$Sci_name_New<- traits$x2[match(catch1$Sci_name, traits$x1)] 
catch.traits$Gregariousness[catch.traits$Sci_name=="Lethrinus mahsena"] <- c(2)
catch.traits$Gregariousness[catch.traits$Sci_name=="Sphyraena jello"] <- c(3)
catch.traits$Gregariousness[catch.traits$Sci_name=="Thunnus albacares"] <- c(3)
catch.traits$Gregariousness[catch.traits$Gregariousness=="#N/A"] <- NA

catch.traits$Gregariousness <- droplevels(as.factor(catch.traits$Gregariousness))

#' collate Trophic groups and give shorter names
tn<-data.frame(TCat=levels(as.factor(catch.traits$TrophicGroup)), short=c("Invert.", "Herb.", "Coral.", "Carni.", "Plankt.", "Herb."))
catch.traits$TGShort <- tn$short[match(catch.traits$TrophicGroup, tn$TCat)]

#' select only Traps, Gillnet and Handline
catch.traits <- subset(catch.traits, gear=="HL" | gear=="TB" | gear=="GN")
catch.traits$ss1 <- as.factor(catch.traits$ss)

#' add number of stations in that area to each line
c3 <- catch.traits %>% dplyr::group_by(ss1) %>% dplyr::summarise( first(depth), first(gear), first(id), first(survey), first(survey_m))
colnames(c3) <- c("ss", "depth", "gear", "id", "survey", "survey_m")
n.area <- c3 %>% dplyr::group_by(id) %>% dplyr::summarise(n_distinct(ss))
colnames(n.area) <- c("id", "nos")

catch.traits$NArea <- n.area$nos[match(catch.traits$id, n.area$id)]
```



## Bathymetric map of Sudan with ggplot and marmap
A nice, bathymetric map of Sudan with management areas overlaid. (Fig. 1 in MS)

March 2021: New code for making bathymetric map to include depth legend, north arrow and scale bar. 
```{r NOAA bathy data, echo=FALSE, warning=FALSE, message=FALSE}
Bathy <- getNOAA.bathy(lon1 = 36.5, lon2 = 39,
                       lat1 = 18, lat2 = 22, resolution = 1)
```


```{r bathymap ggplot, echo=FALSE, warning=FALSE, message=FALSE}
library(mapproj)
library(ggsn)

sudan<-readGEBCO.bathy("./sudan_bathymetry/sudan.nc")

MApoints1<-data.frame(lon=c(37.2, 37.5, 37.37, 37.5, 37.5, 37.8, 38.2), lat=c(21.7, 20.93, 20.3, 19.9, 19, 19.2, 18.5), area=c(1,2,3,4,5,6,7), anames=c("1. \nMarsas \nnorth of \nDungonab", "2. \nDungonab", "3. \nArakiai", "4. \nPort \nSudan", "5. \nSuakin", "6. \nSuakin \narchipelago", "7. Agig"))

cities <- data.frame(lon=c(37.333333, 37.21709967), lat=c(19.1, 19.600512), city=c("Suakin", "Port Sudan"))



#ctd <- data.frame(Longitude = c(37.5, 38, 38.5), Latitude = c(-47, -46.5, -46))

smap2 <- autoplot.bathy(sudan, geom=c("raster")) +
    #scale_fill_gradient2(low="dodgerblue4", mid="white", high="gray10") +
    #geom_point(data = ctd, aes(x = Longitude, y = Latitude), colour = 'black', size = 3, alpha = 1, shape = 15) +
  scale_fill_gradientn(colors = c("midnightblue", "deepskyblue3", "deepskyblue1", "cadetblue3", "cadetblue1", "darkseagreen1", "white", "gray10"),
                            values = scales::rescale(c(-2847, -500, -200, -100, -50, -10, 0, 1, 2010))) +
    geom_path(data=ManageAreas.f,colour="black", aes(x=long, y=lat, group=id)) +
    labs(y = "Latitude", x = "Longitude", fill = "Elevation") 
 
  smap3 <-   smap2 + north(ManageAreas.f, symbol=10,scale=0.2) +geom_text(data=MApoints1, hjust="left", size=3, aes(x=lon, y=lat, group=area, label=anames)) + geom_point(data=cities, size=2, colour="gray10", aes(x=lon, y=lat)) + geom_text(data=cities, label=cities$city, size=4,  hjust=1.15, vjust=0,  aes(x=lon, y=lat, group=city)) +  scalebar(ManageAreas.f, location="topright", dist = 50, dist_unit = "km", transform = TRUE, model = "WGS84") + theme(legend.position = c(0.9, 0.6))

smap3
#ggsave("./figs/Fig1_bathymap2 sudan.tiff")  
  
  
#create world map
world <- map_data("world")
d <- data.frame(x1=c(36.5), x2=c(39), y1=c(18), y2=c(22.3))

regionmap <- ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
    color = "black", fill = "lightgray", size = 0.1
  ) + coord_quickmap(xlim=c(20, 60), ylim=c(0, 40)) + geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), color="red", fill= NA, alpha=0.5) + theme_void() + theme(
  panel.background = element_rect(fill = "white",
                                colour = "white",
                                size = 0.5, linetype = "solid"))
       regionmap
    
#add regionmap as incert to area-map
#png(filename = "./figs/Fig1_bathy_insert.png", width = 3600, height = 2600, res = 300)
tiff(filename = "./figs/Fig1_bathy_insert.tiff", width = 3600, height = 2600, res = 300)

# create a viewport for inset
# vp_inset width/height arguments set the size of the inset; x and y arguments set the position (from 0 to 1) of the left, top corner of the inset along each axis (i.e. not map coordinates as you have in your annotation custom). You can adjust these as you see fit.
vp_inset <- grid::viewport(width = 0.2, height = 0.2, x = 0.29, y = 0.252, just = c("left", "top"))
print(smap3)
print(regionmap, vp = vp_inset)
#while (!is.null(dev.list()))  dev.off()
 dev.off()
```


## Stations plotted on maps, pr survey
Faceted map plotting position of all catch stations for each of the three surveys. 
```{r station map, echo=FALSE, warning=FALSE, message=FALSE}
library(patchwork)

MApoints<-data.frame(lon=c(37.2, 37.53, 37.5, 37.4, 37.3, 37.9, 38.2), lat=c(21.7, 20.92, 20.3, 19.9, 19, 19.3, 18.5), area=c(1,2,3,4,5,6,7), survey_m=c(1,1,1,1,1,1,1))

#ManageAreas.f$survey_m <- 1

world<- map_data("worldHires", c("Sudan", "Ethiopia")) 
world$survey_m<-c("Survey 1: Nov. 2012")
world2<-world
world2$survey_m<-c("Survey 2: May 2013")
world3<-world
world3$survey_m<-c("Survey 3: Nov. 2013")
world<-rbind(world, world2, world3)

colors <- brewer.pal(3, "Dark2")

s1.map <- ggplot(subset(catch, survey==2012901), aes(x=lon, y=lat)) + 
  geom_point(shape="+", size=2, colour= colors[1], aes(group=survey_m, colour= survey_m))  + 
  geom_path(data=ManageAreas.f,colour="dodgerblue3", aes(x=long, y=lat, group=id)) + 
  geom_polygon(data=world, colour="gray35", fill="gray85", aes(x=long, y=lat, group=survey_m))  + 
  coord_cartesian(xlim = c(36.58, 39), ylim=c(17.87, 22.5)) +
  theme_classic() + 
  geom_point(data=ps, size=2, colour="gray35", aes(x=x, y=y, group=survey_m)) + 
  geom_text(data=ps, label="PZU", size=3,  hjust=1, vjust=-1.2,  aes(x=x, y=y, group=survey_m)) + 
  geom_text(data=MApoints, aes(x=lon, y=lat, group=survey_m, label=area)) +
  theme(legend.position="none")  +
  ggsn::north(ManageAreas.f, symbol=10,scale=0.2, location="topright") +  
  ggsn::scalebar(ManageAreas.f, location="bottomleft", dist = 50, dist_unit = "km", transform = TRUE, model = "WGS84", st.size=3, nudge_y=-0.05) + 
  ggtitle("Survey 1: Nov. 2012")


s2.map <- ggplot(subset(catch, survey==2013002), aes(x=lon, y=lat)) + 
  geom_point(shape="+", size=2, colour= colors[2], aes(group=survey_m, colour= survey_m))  + 
  geom_path(data=ManageAreas.f,colour="dodgerblue3", aes(x=long, y=lat, group=id)) + 
  geom_polygon(data=world, colour="gray35", fill="gray85", aes(x=long, y=lat, group=survey_m))  + 
  coord_cartesian(xlim = c(36.58, 39), ylim=c(17.87, 22.5)) +
  theme_classic() + 
  geom_point(data=ps, size=2, colour="gray35", aes(x=x, y=y, group=survey_m)) + 
  geom_text(data=ps, label="PZU", size=3,  hjust=1, vjust=-1.2,  aes(x=x, y=y, group=survey_m)) + 
  geom_text(data=MApoints, aes(x=lon, y=lat, group=survey_m, label=area)) +
  theme(legend.position="none")  + 
  ggtitle("Survey 2: May 2013")


s3.map <- ggplot(subset(catch, survey==2013005), aes(x=lon, y=lat)) + 
  geom_point(shape="+", size=2, colour= colors[3], aes(group=survey_m, colour= survey_m))  + 
  geom_path(data=ManageAreas.f,colour="dodgerblue3", aes(x=long, y=lat, group=id)) + 
  geom_polygon(data=world, colour="gray35", fill="gray85", aes(x=long, y=lat, group=survey_m))  + 
  coord_cartesian(xlim = c(36.58, 39), ylim=c(17.87, 22.5)) +
  theme_classic() + 
  geom_point(data=ps, size=2, colour="gray35", aes(x=x, y=y, group=survey_m)) + 
  geom_text(data=ps, label="PZU", size=3,  hjust=1, vjust=-1.2,  aes(x=x, y=y, group=survey_m)) + 
  geom_text(data=MApoints, aes(x=lon, y=lat, group=survey_m, label=area)) +
  theme(legend.position="none") + 
  ggtitle("Survey 3: Nov. 2013")

s1.map + s2.map +s3.map


ggsave("./figs/Fig2_station_map.tiff", width=240, height=160, units="mm", dpi=400)
ggsave("./figs/Fig2_station_map.png", width=240, height=160, units="mm", dpi=400)
```


## Station information
### Station table
Table describing the sampling effort in each area pr survey, number of different gear types, max, min and average depth, number of traps with / without catches. 
```{r station table, echo=FALSE, warning=FALSE, message=FALSE}
#' TABLE OF SAMPLING EFFORT PR. MANAGEMENT AREA
#' (new for resubmission to PlosOne)
#' --------------------------------------------

s <- unique(catch.unmod$survey)
a <- as.factor(1:7)
ay_info <- data.frame(survey=integer(), id=integer(), Ntraps=integer(), Nhl=integer(), NGn=integer(), TBhrs=double(), HLhrs=double(), GNhrs=double(), DepthAvg=double(), DepthSD=double(),DepthMax=double(), DepthMin=double()) 
cn <- colnames(ay_info)
# depth for TB only (GN are at the surface, and HL not measured)

for (i in 1:length(s)) { 
  y <- subset(catch.unmod, survey==s[i])
  for (j in 1: length(a)){
    ay_1 <- data.frame(survey=integer(), id=integer(), Ntraps=integer(), Nhl=integer(), NGn=integer(), TBhrs=double(), HLhrs=double(), GNhrs=double(), DepthAvg=double(), DepthSD=double(), DepthMax=double(), DepthMin=double())
    ya <- subset(y, id==a[j])
    ay_1 <- rbind(ay_1, c(
      as.integer(as.character(ya$survey[1])), 
      as.integer(as.character(ya$id[1])), 
      length(unique(subset(ya, gear=="TB")$station)),  
      length(unique(subset(ya, gear=="HL")$station)), 
      length(unique(subset(ya, gear=="GN")$station)), 
      sum(subset(ya, gear=="TB")[!duplicated(ya$station),]$Fhrs, na.rm=TRUE),  
      sum(subset(ya, gear=="HL")[!duplicated(ya$station),]$Fhrs, na.rm=TRUE), 
      sum(subset(ya, gear=="GN")[!duplicated(ya$station),]$Fhrs, na.rm=TRUE), 
      mean(subset(ya, gear=="TB")[!duplicated(ya$station),]$depth, na.rm=TRUE), 
      sd(subset(ya, gear=="TB")[!duplicated(ya$station),]$depth, na.rm=TRUE),
      max(subset(ya, gear=="TB")$depth, na.rm=TRUE), 
      min(subset(ya, gear=="TB")$depth, na.rm=TRUE)    ))
    colnames(ay_1) <- cn
    ay_info <- rbind(ay_info, ay_1)
    colnames(ay_info) <- cn
  }
}

formattable(ay_info)

write.csv2(ay_info, "./figs/Area_Station_table.csv")
```

#### Distance between stations
```{r station distance, echo=FALSE, warning=FALSE, message=FALSE}
library(sf)

#### TRAPS
sts <- subset(catch.unmod, gear=="TB")[,c(1,30, 17, 13,14)]

# split by survey and area
dist_tbl <- data.frame(dist=double(), survey=integer(), area=integer())
surveys <- c(2012901, 2013002, 2013005)
areas <- c(1:7)

for (i in 1:3) {   
  for (j in 1:7) { 
    sts2 <- subset(sts, survey==surveys[i] & id ==areas[j])
    s_sf<- sts2 %>% 
      st_as_sf(coords = c("lon", "lat")) %>%
      st_set_crs(4326) #'4326' is same as WGS84
    dist2 <- as.data.frame(st_distance(s_sf))
    
    d2 <- data.frame(dist = c(t(dist2)))
    d2$survey <- surveys[i]
    d2$area <- areas[j]
    d2 <- subset(d2, dist!=0)
    
    dist_tbl <- rbind(dist_tbl, d2)
  }
}
#plot and facet_wrap by surveys

survey_names <- c(`2012901` = "Nov. 2012", `2013002` = "May 2013", `2013005` = "Nov. 2013" )

dist.plot <- ggplot(dist_tbl, aes(x=as.factor(area), y=dist/1000, group=as.factor(area))) + geom_boxplot() + facet_wrap(vars(survey), labeller = as_labeller(survey_names)) +ylab("Distance between stations (km)") +xlab("Management areas")
dist.plot

ggsave("./figs/S1 Traps_station_distances.png")

# table of stats
dist_tbl %>% group_by(survey, area) %>% summarise( min=min(dist),max=max(dist), mean=mean(dist), median=median(dist))%>% write.csv2("./figs/disttable_traps.csv")



#### GILLNETS
sts <- subset(catch.unmod, gear=="GN")[,c(1,30, 17, 13,14)]

# split by survey and area
dist_tbl <- data.frame(dist=double(), survey=integer(), area=integer())
surveys <- c(2012901, 2013002, 2013005)
areas <- c(1:7)

for (i in 1:3) {   
  for (j in 1:7) { 
    sts2 <- subset(sts, survey==surveys[i] & id ==areas[j])
    if (nrow(sts2)>0) { 
      s_sf<- sts2 %>% 
        st_as_sf(coords = c("lon", "lat")) %>%
        st_set_crs(4326) #'4326' is same as WGS84
      dist2 <- as.data.frame(st_distance(s_sf))
      
      d2 <- data.frame(dist = c(t(dist2)))
      d2$survey <- surveys[i]
      d2$area <- areas[j]
      d2 <- subset(d2, dist!=0)
      
      dist_tbl <- rbind(dist_tbl, d2)
    }
  }
}
#plot and facet_wrap by surveys

survey_names <- c(`2012901` = "Nov. 2012", `2013002` = "May 2013", `2013005` = "Nov. 2013" )

dist.plot <- ggplot(dist_tbl, aes(x=as.factor(area), y=dist/1000, group=as.factor(area))) + geom_boxplot() + facet_wrap(vars(survey), labeller = as_labeller(survey_names)) +ylab("Distance between stations (km)") +xlab("Management areas")
dist.plot

#ggsave("./figs/Gillnets_station_distances.png")

# table of stats
dist_tbl %>% group_by(survey, area) %>% summarise( min=min(dist),max=max(dist), mean=mean(dist), median=median(dist)) %>% write.csv2("./figs/disttable_nets.csv")
```



#### Species table
Number of fish (organized by family and species) caught by Gillnet or Traps for each survey, and in total across all surveys and gears. 

* (new table for revision 2020) *
```{r species tabel, echo=FALSE, warning=FALSE, message=FALSE}

species.table <- subset(catch, Sci_name!="NO CATCH" & Sci_name!="No catch" & gear=="TB" | gear =="GN") %>% dplyr::group_by(Fam_name, Sci_name, survey, gear) %>% dplyr::summarise(No.individ=sum(as.integer(number), No.gear=n_distinct(ss)))

#number of species and families total
subset(catch, Sci_name!="NO CATCH" & Sci_name!="No catch" & gear=="TB" | gear =="GN") %>% summarise(families=n_distinct(Fam_name), species=n_distinct(Sci_name))

#number of species and families by area
species.area.table <-  subset(catch, Sci_name!="NO CATCH" & Sci_name!="No catch" & gear=="TB" | gear =="GN") %>% 
  group_by(id, gear) %>%
  summarise( species=n_distinct(Sci_name)) %>%
  pivot_wider(names_from = gear, values_from = species)


#number of species and families by gear
species.table %>% dplyr::group_by(gear) %>% dplyr::summarise(families=n_distinct(Fam_name), species=n_distinct(Sci_name))

gnsp <- subset(catch, Sci_name!="NO CATCH" & Sci_name!="No catch" &  gear =="GN") %>% dplyr::group_by(Fam_name, Sci_name) %>% dplyr::summarise(No.individ=sum(as.integer(number), No.gear=n_distinct(ss)))

species.table <- spread(species.table, gear, No.individ)

spt <- species.table[,2:5] %>%
  gather(variable, value, -(Sci_name:survey)) %>%
  unite(temp, survey, variable) %>%
  spread(temp, value)  %>%
  replace(is.na(.), 0) %>%
  rowwise() %>%
  #mutate(sum = rowSums(.[2:7]))
  dplyr::mutate(sum = sum(c_across("2012901_GN":"2013005_TB")))

spt$fam_name <- species.table$Fam_name[match(spt$Sci_name, species.table$Sci_name)] 
 
spt[,c(9,1:8)]%>%
  kable(format = "html", escape = F) %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 8)

write.csv2(spt, "./figs/species_catch_table.csv")


# similar table for number of traps/gillnets with catches of species

st2 <- subset(catch, Sci_name!="NO CATCH" & Sci_name!="No catch" & gear=="TB" | gear =="GN") %>% dplyr::group_by(Fam_name, Sci_name, survey, gear) %>% dplyr::summarise( No.gear=n_distinct(ss))

st2 <- spread(st2, gear, No.gear)

spt2 <- st2[,2:5] %>%
  gather(variable, value, -(Sci_name:survey)) %>%
  unite(temp, survey, variable) %>%
  spread(temp, value)  %>%
  replace(is.na(.), 0) %>%
  rowwise() %>%
  #mutate(sum = rowSums(.[2:7]))
  dplyr::mutate(sum = sum(c_across("2012901_GN":"2013005_TB")))

spt2$fam_name <- species.table$Fam_name[match(spt$Sci_name, species.table$Sci_name)] 
 
spt2[,c(9,1:8)]%>%
  kable(format = "html", escape = F) %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 8)

write.csv2(spt2, "./figs/species_catch_table_NoStations.csv")


```


***

### Number of set traps per depth
Box plot of depths at wich traps were set. I prefer this visualization of depth ranges of the traps, rather than a (boring) table. 
```{r traps per depth, echo=FALSE, warning=FALSE, message=FALSE}

# New facet label names for survey variable
#survey.labs <- c("Nov.2012", "May.2013", "Nov.2013")
#names(survey.labs) <- c("Survey 1: Nov. 2012", "Survey 2: May 2013",  "Survey 3: Nov. 2013")

c2 <- catch %>% dplyr::group_by(ss) %>% dplyr::summarise(first(depth), first(gear), first(id), first(survey), first(survey_m))
colnames(c2) <- c("ss", "depth", "gear", "id", "survey", "survey_m")
c2$survey_m <- as.factor(c2$survey_m)



# function for number of observations 
give.n <- function(x){
  return(c(y = median(x)*0.7, label = length(x))) 
  # experiment with the multiplier to find the perfect position
}

dp <- ggplot(subset(c2, gear=="TB" & depth!=0), aes(as.factor(id), depth))

#dp <- dp +  geom_violin(scale="width", aes(fill=as.factor(id)), draw_quantiles = c(0.5)) + scale_y_reverse() +stat_summary(fun.data = give.n, geom = "text", fun.y = median) + labs(x="Management areas", y="Trap set depth (m)") + scale_fill_manual(name="Management area",values=sequential_hcl(n = 7, h = c(0, -100), c = c(80, NA, 40), l = c(53, 75), power = c(1, 1), rev = TRUE, register = ), guide=FALSE ) +theme_bw()  +facet_wrap(~survey_m, labeller=labeller(survey_m=c("Survey 1: Nov. 2012" = "Nov.2012",      "Survey 2: May 2013" = "May.2013",      "Survey 3: Nov. 2013" = "Nov.2013")))

dp <- dp +  geom_boxplot(scale="width", aes(), draw_quantiles = c(0.5)) + scale_y_reverse() + labs(x="Management areas", y="Trap set depth (m)")  +theme_bw()  +facet_wrap(~survey_m, labeller=labeller(survey_m=c("Survey 1: Nov. 2012" = "Nov.2012",      "Survey 2: May 2013" = "May.2013",      "Survey 3: Nov. 2013" = "Nov.2013")))

dp

ggsave("./figs/trap_depth.tiff", width=6, height=6)
ggsave("./figs/trap_depth.png", width=6, height=6)

#tiff(file="./figs/trap_area_depth.tiff", width=3000, height=1500, res=400, pointsize=10, compression=c("none"))
#dp
#dev.off()
```

####Test differneces in depths in each area between years and between areas

```{r  test trap depth, echo=FALSE, warning=FALSE, message=FALSE}

c3 <- subset(catch, gear=="TB") %>% 
  dplyr::group_by(ss) %>% 
  dplyr::summarise(id=first(id), Fhrs=first(Fhrs), depth=first(depth), gear=first(gear), id=first(id), survey=first(survey), survey_m=first(survey_m))

c3$sa <- paste(substring(c3$survey_m, 11,13), substring(c3$survey, 3,4),  "A", c3$id, sep="")

#' test for normality
shapiro.test(c3$depth)

# significan difference, e.g null-hypothesis that data are normally distributed is rejected

#' Q-Q plots
ggplot(data = c3, mapping = aes(sample = depth,  fill = factor(id))) + qqplotr::stat_qq_band() + qqplotr::stat_qq_line() + qqplotr::stat_qq_point(size=0.2) + facet_wrap(~ id, scales="free") +  labs(x = "depth predicted", y = "depth Observed") + ggtitle("Q-Q plots depth")

#CONCLUSION: Depth is NOT normally distributed, so can use ANOVA for analysis of variance

# Use non-parametric Kruskal-Wallis test

#' 1 - test for difference between areas
kruskal.test(c3$depth, c3$sa)
#significant difference

#' Post-Hoc tests of Kruskal - Wallis results

#' using Conover - Iman test as this is described as more powerful than other comparisons (Wikipedia Kruskal-Wallis)
#' using 'holm' (Holm-BOnferroni correction) as this is more conservative (or as) as Bonferroni method
ct <- conover.test::conover.test(c3$depth, as.factor(c3$sa), method="holm")

#convert Conover test output into nice looking table

l1<-numeric(21)

l1[1] <- 1
for (i in 2:21) l1[i] <- l1[i-1] +(i-1)
l2 <- c(1:20)

df0 <- matrix(0, nrow = 21, ncol = 21)
#df0[1,1] <- ct$P.adjusted[1]

for (i in 2:21){
  for (j in 1:l2[i-1]) { 
    df0[i,j] <- ct$P.adjusted[l1[i-1]+j-1]
  }
}


df0 <- as.data.frame(df0)
colnames(df0) <- levels(as.factor(c3$sa))
rownames(df0) <- levels(as.factor(c3$sa))
df0 <- round(df0,5)

df0[df0==0]<-" "
df0[df0==1] <- "1.00000"

write.csv2(df0, "trap depth K_W test.csv")

df0%>%
  tibble::rownames_to_column('RN') %>%
  dplyr::mutate(
    May13A1 = cell_spec(May13A1, "html", color = ifelse(May13A1 <= 0.05, "red", "black")),
    May13A2 = cell_spec(May13A2, "html", color = ifelse(May13A2 <= 0.05, "red", "black")),
    May13A3 = cell_spec(May13A3, "html", color = ifelse(May13A3 <= 0.05, "red", "black")),
    May13A4 = cell_spec(May13A4, "html", color = ifelse(May13A4 <= 0.05, "red", "black")),
    May13A5 = cell_spec(May13A5, "html", color = ifelse(May13A5 <= 0.05, "red", "black")),
    May13A6 = cell_spec(May13A6, "html", color = ifelse(May13A6 <= 0.05, "red", "black")),
    May13A7 = cell_spec(May13A7, "html", color = ifelse(May13A7 <= 0.05, "red", "black")),
    Nov12A1 = cell_spec(Nov12A1, "html", color = ifelse(Nov12A1 <= 0.05, "red", "black")),
    Nov12A2 = cell_spec(Nov12A2, "html", color = ifelse(Nov12A2 <= 0.05, "red", "black")),
    Nov12A3 = cell_spec(Nov12A3, "html", color = ifelse(Nov12A3 <= 0.05, "red", "black")),
    Nov12A4 = cell_spec(Nov12A4, "html", color = ifelse(Nov12A4 <= 0.05, "red", "black")),
    Nov12A5 = cell_spec(Nov12A5, "html", color = ifelse(Nov12A5 <= 0.05, "red", "black")),
    Nov12A6 = cell_spec(Nov12A6, "html", color = ifelse(Nov12A6 <= 0.05, "red", "black")),
    Nov12A7 = cell_spec(Nov12A7, "html", color = ifelse(Nov12A7 <= 0.05, "red", "black")),
    Nov13A1 = cell_spec(Nov13A1, "html", color = ifelse(Nov13A1 <= 0.05, "red", "black")),
    Nov13A2 = cell_spec(Nov13A2, "html", color = ifelse(Nov13A2 <= 0.05, "red", "black")),
    Nov13A3 = cell_spec(Nov13A3, "html", color = ifelse(Nov13A3 <= 0.05, "red", "black")),
    Nov13A4 = cell_spec(Nov13A4, "html", color = ifelse(Nov13A4 <= 0.05, "red", "black")),
    Nov13A5 = cell_spec(Nov13A5, "html", color = ifelse(Nov13A5 <= 0.05, "red", "black")),
    Nov13A6 = cell_spec(Nov13A6, "html", color = ifelse(Nov13A6 <= 0.05, "red", "black")),
    Nov13A7 = cell_spec(Nov13A7, "html", color = ifelse(Nov13A7 <= 0.05, "red", "black")) )%>%
  tibble::column_to_rownames('RN') %>%
  kable(format = "html", escape = F, rownames=TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
   add_header_above(c(" " = 1, "May 2013" = 7, "Nov 2012" = 7, "Nov 2013" = 7))
  


```
Depth data were not normally distributed and hence the difference in depth could only be analyzed using non-parametric Kruskal - Wallis rank sum test, and Conover - Iman post hoc test with Holm-Bonferroni correction. 

Of the 210 survey - area comparisons, depth distributions were significantly different between areas as follows:
  
May13A5 - Nov12A2  
May13A5 - Nov13A2  
Nov12A2 - Nov12A5  
Nov12A5 - Nov13A2  
Nov12A2 - Nov13A5  
  
Which corresponds with the plot of depth-distributions by survey and area above. 



*** 

### Duration of trap sets 

#### Mean, min & max hrs. of fishing
```{r hrs of fishing, echo=FALSE, warning=FALSE, message=FALSE}
subset(catch, gear=="TB") %>% 
  dplyr::group_by(ss) %>% 
  dplyr::summarise(id=first(id), fhrs=first(Fhrs)) %>% 
  dplyr::group_by(id) %>%
  dplyr::summarise( maxhrs=max(fhrs), minhrs=min(fhrs), meanhrs=mean(fhrs)) %>%
  kable(format = "html", escape = F) %>%
  kable_styling("striped", full_width = F)

#overall mean Fhrs
summary(subset(catch, gear=="GN") %>% 
  group_by(ss) %>% 
  dplyr::summarise(id=first(id), fhrs=first(Fhrs)))

```


### Fishing hours (traps) by survey and area
Boxplot of median with mean values plotted as red circle. 
```{r duration traps, echo=FALSE, warning=FALSE, message=FALSE}

c3 <- subset(catch, gear=="TB") %>% 
  dplyr::group_by(ss) %>% 
  dplyr::summarise(id=first(id), Fhrs=first(Fhrs), depth=first(depth), gear=first(gear), id=first(id), survey=first(survey), survey_m=first(survey_m))
c3$survey_m <- as.factor(c3$survey_m)


duration.plot <- ggplot(c3, aes(as.factor(id), Fhrs))

duration.plot <-  duration.plot +  geom_boxplot() + stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="red", fill="red") + labs(x="Management areas", y="Hours fishing (traps)")  +theme_bw() +facet_wrap(~survey_m, labeller=labeller(survey_m=c("Survey 1: Nov. 2012" = "Nov.2012",
      "Survey 2: May 2013" = "May.2013",
      "Survey 3: Nov. 2013" = "Nov.2013")))

duration.plot
#a1 <- subset(c3, id==1)
#ggplot(a1, aes(as.factor(id), Fhrs)) + geom_boxplot()+facet_wrap(~as.factor(survey_m))

#by(a1$Fhrs, a1$survey, summary)
```
The average trap soak-time (fishing hours) across all surveys and areas was 17.984 hrs with a median of 16.15, but the November 2012 survey had more varied and higher soak times than the May 2013 and Nov. 2013. 

#### Test differences in soak times (traps) K-W test
```{r  test Fhrs traps, echo=FALSE, warning=FALSE, message=FALSE}
#' test for normality
shapiro.test(c3$Fhrs)

#' Q-Q plots
ggplot(data = c3, mapping = aes(sample = Fhrs,  fill = factor(id))) + qqplotr::stat_qq_band() + qqplotr::stat_qq_line() + qqplotr::stat_qq_point(size=0.2) + facet_wrap(~ id, scales="free") +  labs(x = "Fhrs predicted", y = "Fhrs Observed") + ggtitle("Q-Q plots depth")

c3$sa <- paste(substring(c3$survey_m, 11,13), substring(c3$survey, 3,4),  "A", c3$id, sep="")
#CONCLUSION: Fishing hours (soak time) is NOT normally distributed, so can use ANOVA for analysis of variance

# Use non-parametric Kruskal-Wallis test

#' 1 - test for difference between areas
kruskal.test(c3$Fhrs, c3$sa)
#significant difference

#' Post-Hoc tests of Kruskal - Wallis results
#' using Conover - Iman test as this is described as more powerful than other comparisons (Wikipedia Kruskal-Wallis)
#' using 'holm' (Holm-BOnferroni correction) as this is more conservative (or as) as Bonferroni method

ct <- conover.test::conover.test(c3$Fhrs, as.factor(c3$sa), method="holm")

#convert Conover test output into nice looking table

l1<-numeric(21)

l1[1] <- 1
for (i in 2:21) l1[i] <- l1[i-1] +(i-1)
l2 <- c(1:20)

df0 <- matrix(0, nrow = 21, ncol = 21)
#df0[1,1] <- ct$P.adjusted[1]

for (i in 2:21){
  for (j in 1:l2[i-1]) { 
    df0[i,j] <- ct$P.adjusted[l1[i-1]+j-1]
  }
}


df0 <- as.data.frame(df0)
colnames(df0) <- levels(as.factor(c3$sa))
rownames(df0) <- levels(as.factor(c3$sa))
#df0 <- round(df0,5)

#df0[df0==0]<-" "
#df0[df0==1] <- "1.00000"

write.csv2(df0, "trap soak time K_W test.csv")

df0%>%
  tibble::rownames_to_column('RN') %>%
  dplyr::mutate(
    May13A1 = cell_spec(May13A1, "html", color = ifelse(May13A1 <= 0.05, "red", "black")),
    May13A2 = cell_spec(May13A2, "html", color = ifelse(May13A2 <= 0.05, "red", "black")),
    May13A3 = cell_spec(May13A3, "html", color = ifelse(May13A3 <= 0.05, "red", "black")),
    May13A4 = cell_spec(May13A4, "html", color = ifelse(May13A4 <= 0.05, "red", "black")),
    May13A5 = cell_spec(May13A5, "html", color = ifelse(May13A5 <= 0.05, "red", "black")),
    May13A6 = cell_spec(May13A6, "html", color = ifelse(May13A6 <= 0.05, "red", "black")),
    May13A7 = cell_spec(May13A7, "html", color = ifelse(May13A7 <= 0.05, "red", "black")),
    Nov12A1 = cell_spec(Nov12A1, "html", color = ifelse(Nov12A1 <= 0.05, "red", "black")),
    Nov12A2 = cell_spec(Nov12A2, "html", color = ifelse(Nov12A2 <= 0.05, "red", "black")),
    Nov12A3 = cell_spec(Nov12A3, "html", color = ifelse(Nov12A3 <= 0.05, "red", "black")),
    Nov12A4 = cell_spec(Nov12A4, "html", color = ifelse(Nov12A4 <= 0.05, "red", "black")),
    Nov12A5 = cell_spec(Nov12A5, "html", color = ifelse(Nov12A5 <= 0.05, "red", "black")),
    Nov12A6 = cell_spec(Nov12A6, "html", color = ifelse(Nov12A6 <= 0.05, "red", "black")),
    Nov12A7 = cell_spec(Nov12A7, "html", color = ifelse(Nov12A7 <= 0.05, "red", "black")),
    Nov13A1 = cell_spec(Nov13A1, "html", color = ifelse(Nov13A1 <= 0.05, "red", "black")),
    Nov13A2 = cell_spec(Nov13A2, "html", color = ifelse(Nov13A2 <= 0.05, "red", "black")),
    Nov13A3 = cell_spec(Nov13A3, "html", color = ifelse(Nov13A3 <= 0.05, "red", "black")),
    Nov13A4 = cell_spec(Nov13A4, "html", color = ifelse(Nov13A4 <= 0.05, "red", "black")),
    Nov13A5 = cell_spec(Nov13A5, "html", color = ifelse(Nov13A5 <= 0.05, "red", "black")),
    Nov13A6 = cell_spec(Nov13A6, "html", color = ifelse(Nov13A6 <= 0.05, "red", "black")),
    Nov13A7 = cell_spec(Nov13A7, "html", color = ifelse(Nov13A7 <= 0.05, "red", "black")) )%>%
  tibble::column_to_rownames('RN') %>%
  kable(format = "html", escape = F, rownames=TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
   add_header_above(c(" " = 1, "May 2013" = 7, "Nov 2012" = 7, "Nov 2013" = 7))
  

```
Trap soak times:  
Of the 210 area- survey combinations 69 were significantly different (Kruskal Wallis rank sum test with Conover-Iman post-hoc test w Holm correction, p<0.05):

May13A1 - May13A6  
May13A1 - Nov12A1  
May13A1 - Nov12A2  
May13A1 - Nov12A5  
May13A1 - Nov12A6  
May13A1 - Nov12A7  
  
May13A2 - May13A6  
May13A2 - May13A7  
May13A2 - Nov12A1  
May13A2 - Nov12A2  
May13A2 - Nov12A5  
May13A2 - Nov12A6  
May13A2 - Nov12A7  
  
May13A3 - Nov12A6  
May13A3 - Nov12A7  
  
May13A4 - Nov12A1  
May13A4 - Nov12A5  
May13A4 - Nov12A6  
May13A4 - Nov12A7  
  
May13A5 - Nov12A5  
May13A5 - Nov12A6  
May13A5 - Nov12A7  
  
May13A6 - Nov12A6  
May13A6 - Nov12A7  
May13A6 - Nov13A2  
May13A6 - Nov13A3  
May13A6 - Nov13A5  
  
May13A7 - Nov12A6  
May13A7 - Nov12A7  
May13A7 - Nov13A2  
May13A7 - Nov13A5  
  
Nov12A1 - Nov12A4  
Nov12A1 - Nov13A2  
Nov12A1 - Nov13A3  
Nov12A1 - Nov13A4  
Nov12A1 - Nov13A5  
Nov12A1 - Nov13A6  
  
Nov12A2 - Nov12A6  
Nov12A2 - Nov12A7  
Nov12A2 - Nov13A2  
Nov12A2 - Nov13A3  
Nov12A2 - Nov13A5  
  
Nov12A3 - Nov12A6  
Nov12A3 - Nov12A7  
Nov12A3 - Nov13A5  
  
Nov12A4 - Nov12A5  
Nov12A4 - Nov12A6  
Nov12A4 - Nov12A7  
  
Nov12A5 - Nov13A2  
Nov12A5 - Nov13A3  
Nov12A5 - Nov13A4  
Nov12A5 - Nov13A5  
Nov12A5 - Nov13A6  
  
Nov12A6 - Nov13A1  
Nov12A6 - Nov13A2  
Nov12A6 - Nov13A3  
Nov12A6 - Nov13A4  
Nov12A6 - Nov13A5  
Nov12A6 - Nov13A6  
Nov12A6 - Nov13A7  
  
Nov12A7 - Nov13A1  
Nov12A7 - Nov13A2  
Nov12A7 - Nov13A3  
Nov12A7 - Nov13A4  
Nov12A7 - Nov13A5  
Nov12A7 - Nov13A6  
Nov12A7 - Nov13A7  
  
Nov13A1 - Nov13A7  
  
Nov13A5 - Nov13A7  


eys. 


### Duration and catch rates of gill-net sets
```{r gillnet set time, echo=FALSE, warning=FALSE, message=FALSE}

c33 <- subset(catch, gear=="GN") %>% 
  dplyr::group_by(ss) %>% 
  dplyr::summarise(id=first(id), Fhrs=first(Fhrs), depth=first(depth), gear=first(gear), id=first(id), survey=first(survey), survey_m=first(survey_m))
c33$survey_m <- as.factor(c33$survey_m)


#duration.plot.gn <- ggplot(c33, aes(id, Fhrs))

#duration.plot.gn <-  duration.plot.gn +  geom_point() + stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="red", fill="red") + labs( x="Management areas", y="Hours fishing (gillnets)")  +theme_bw() +facet_wrap(~survey_m, labeller=labeller(survey_m=c("Survey 1: Nov. 2012" = "Nov.2012",      "Survey 2: May 2013" = "May.2013",      "Survey 3: Nov. 2013" = "Nov.2013")))

duration.plot.gn <- ggplot(c33, aes(as.factor(id), Fhrs))

duration.plot.gn <- duration.plot.gn  + geom_boxplot()  + stat_summary(fun.y=mean, geom="point", shape=20, size=3, color="red", fill="red") + labs( x="Management areas", y="Hours fishing (gillnets)")  +theme_bw() +facet_wrap(~survey_m, labeller=labeller(survey_m=c("Survey 1: Nov. 2012" = "Nov.2012",
      "Survey 2: May 2013" = "May.2013",
      "Survey 3: Nov. 2013" = "Nov.2013")))

duration.plot.gn

# catch rates for gillnets
cpue.gn <- subset(catch.cpue, gear=="GN") %>% 
  dplyr::group_by(ss) %>% 
  dplyr::summarise(id=first(id), Fhrs=first(Fhrs), CPUEw=sum(CPUEw), depth=first(depth), gear=first(gear), id=first(id), survey=first(survey), survey_m=first(survey_m))
cpue.gn$survey_m <- as.factor(cpue.gn$survey_m)

ca.plot2<-ggplot(cpue.gn, aes(as.factor(id), CPUEw))


ca.plot2 <- ca.plot2 + geom_boxplot()+ labs( x="Management areas", y="CPUE kg/hour fishing (gillnets)") +theme_bw() +facet_wrap(~survey_m, labeller=labeller(survey_m=c("Survey 1: Nov. 2012" = "Nov.2012",
      "Survey 2: May 2013" = "May.2013",
      "Survey 3: Nov. 2013" = "Nov.2013")))
ca.plot2



#(duration.plot.gn | ca.plot2) + plot_annotation(tag_levels = 'A') & 
 # theme(plot.tag = element_text(size = 10)) 

```
Fishing hours for traps was significantly affected by areas (except area 3 which had a very large variability), however, the number of gillnet sets was very low pr area and survey. 

#### Test of difference in fishing hours, gillnets
```{r gillnet fishing hours K-W test, echo=FALSE, warnings=FALSE, message=FALSE}

c33$sa <-  paste(substring(c33$survey_m, 11,13), substring(c33$survey, 3,4),  "A", c33$id, sep="")

ct <- conover.test::conover.test(c33$Fhrs, as.factor(c33$sa), method="holm")

#convert Conover test output into nice looking table

l1<-numeric(18)

l1[1] <- 1
for (i in 2:18) l1[i] <- l1[i-1] +(i-1)
l2 <- c(1:18)

df0 <- matrix(0, nrow = 18, ncol = 18)
#df0[1,1] <- ct$P.adjusted[1]

for (i in 2:18){
  for (j in 1:l2[i-1]) { 
    df0[i,j] <- ct$P.adjusted[l1[i-1]+j-1]
  }
}


df0 <- as.data.frame(df0)
colnames(df0) <- levels(as.factor(c33$sa))
rownames(df0) <- levels(as.factor(c33$sa))
df0 <- round(df0,5)

df0[df0==0]<-" "
df0[df0==1] <- "1.00000"

write.csv2(df0, "gill nets soak time K_W test.csv")

df0%>%
  tibble::rownames_to_column('RN') %>%
  dplyr::mutate(
    May13A1 = cell_spec(May13A1, "html", color = ifelse(May13A1 <= 0.05, "red", "black")),
    May13A2 = cell_spec(May13A2, "html", color = ifelse(May13A2 <= 0.05, "red", "black")),
    May13A3 = cell_spec(May13A3, "html", color = ifelse(May13A3 <= 0.05, "red", "black")),
    May13A4 = cell_spec(May13A4, "html", color = ifelse(May13A4 <= 0.05, "red", "black")),
    May13A5 = cell_spec(May13A5, "html", color = ifelse(May13A5 <= 0.05, "red", "black")),
    May13A6 = cell_spec(May13A6, "html", color = ifelse(May13A6 <= 0.05, "red", "black")),
    Nov12A2 = cell_spec(Nov12A2, "html", color = ifelse(Nov12A2 <= 0.05, "red", "black")),
    Nov12A3 = cell_spec(Nov12A3, "html", color = ifelse(Nov12A3 <= 0.05, "red", "black")),
    Nov12A5 = cell_spec(Nov12A5, "html", color = ifelse(Nov12A5 <= 0.05, "red", "black")),
    Nov12A6 = cell_spec(Nov12A6, "html", color = ifelse(Nov12A6 <= 0.05, "red", "black")),
    Nov12A7 = cell_spec(Nov12A7, "html", color = ifelse(Nov12A7 <= 0.05, "red", "black")),
    Nov13A1 = cell_spec(Nov13A1, "html", color = ifelse(Nov13A1 <= 0.05, "red", "black")),
    Nov13A2 = cell_spec(Nov13A2, "html", color = ifelse(Nov13A2 <= 0.05, "red", "black")),
    Nov13A3 = cell_spec(Nov13A3, "html", color = ifelse(Nov13A3 <= 0.05, "red", "black")),
    Nov13A4 = cell_spec(Nov13A4, "html", color = ifelse(Nov13A4 <= 0.05, "red", "black")),
    Nov13A5 = cell_spec(Nov13A5, "html", color = ifelse(Nov13A5 <= 0.05, "red", "black")),
    Nov13A6 = cell_spec(Nov13A6, "html", color = ifelse(Nov13A6 <= 0.05, "red", "black")),
    Nov13A7 = cell_spec(Nov13A7, "html", color = ifelse(Nov13A7 <= 0.05, "red", "black")) )%>%
  tibble::column_to_rownames('RN') %>%
  kable(format = "html", escape = F, rownames=TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
   add_header_above(c(" " = 1, "May 2013" = 6, "Nov 2012" = 5, "Nov 2013" = 7))
  


```
Gill net fishing time:  
K-W test significant, but only 2 were significant (Conover-Iman post-hoc test w Holm correction, p<0.05):

Nov13A1 - May13A2  
Nov13A1 - May13A5  

*** 

## Analysis of catches by traits and other factors

### Catch rates pr area & survey
Must aggregate data per station
```{r CPUE traps area, echo=FALSE, warning=FALSE, message=FALSE}

cpue.tr.st <- subset(catch.cpue, gear=="TB") %>% 
  dplyr::group_by(ss) %>% 
  dplyr::summarise(id=first(id), Fhrs=first(Fhrs), CPUEw=sum(CPUEw), depth=first(depth), gear=first(gear), id=first(id), survey=first(survey), survey_m=first(survey_m))
cpue.tr.st$survey_m <- as.factor(cpue.tr.st$survey_m)

ca.plot<-ggplot(subset(cpue.tr.st,  depth!=0), aes(as.factor(id), CPUEw))


#ca.plot <- ca.plot + geom_violin(scale="width", aes(fill=as.factor(id)), draw_quantiles = c(0.5))  + labs(x="Management areas", y="CPUE kg/hour fishing (traps") + scale_fill_manual(name="Management area",values=sequential_hcl(n = 7, h = c(0, -100), c = c(80, NA, 40), l = c(53, 75), power = c(1, 1), rev = TRUE, register = ), guide=FALSE ) +theme_bw()+facet_wrap(~survey_m, labeller=labeller(survey_m=c("Survey 1: Nov. 2012" = "Nov.2012",      "Survey 2: May 2013" = "May.2013",      "Survey 3: Nov. 2013" = "Nov.2013")))

ca.plot <- ca.plot + geom_boxplot(scale="width", aes( draw_quantiles = c(0.5)))  + labs(x="Management areas", y="CPUE kg/hour fishing (traps") + scale_fill_manual(name="Management area",values=sequential_hcl(n = 7, h = c(0, -100), c = c(80, NA, 40), l = c(53, 75), power = c(1, 1), rev = TRUE, register = ), guide=FALSE ) +theme_bw()+facet_wrap(~survey_m, labeller=labeller(survey_m=c("Survey 1: Nov. 2012" = "Nov.2012",      "Survey 2: May 2013" = "May.2013",      "Survey 3: Nov. 2013" = "Nov.2013")))

ca.plot

#tiff(file="./figs/catch_rates_traps_area_survey.tiff", width=3000, height=2000, res=400, pointsize=10, compression=c("none"))
#ca.plot
#dev.off()
```


## Statistical analysis CPUE catches
### Testing for normality of CPUEw of trap catches
Shapiro test for normality and Q-Q plots shows that data is non-normal and zero-inflated. 

```{r normality test, echo=FALSE, warning=FALSE, message=FALSE}

#' test for normality
shapiro.test(cpue.tr.st$CPUEw)

#' Q-Q plots
library("qqplotr")
gg <- ggplot(data = cpue.tr.st, mapping = aes(sample = CPUEw,  fill = factor(id))) + stat_qq_band() + stat_qq_line() + stat_qq_point(size=0.2) + facet_wrap(~ id, scales="free") +  labs(x = "CPUEw predicted", y = "CPUEw Observed") + ggtitle("Q-Q plots CPUEw")
gg
```
### Test for difference in CPUEw of traps between areas (non-parametric test)

CPUE data are non-normal (see Shapiro-test & Q-Q plots above) & zero-inflated (many stations with 0 catch). 

**Zero-inflated GAM model**
A GAM model approach is more appropriate than a non-parametric approach (e.g Kruskal Wallis)  as it evaluates the effects of all factors at the same time, instead of evaluating the effects of area separate from survey. 

* Developed seperate trap and gillnet models for CPUE (weight and numbers) including depth, area and survey. 

* Tested models where survey was included as a random factor, although this excludes evaluation of the potential significant effects of the surveys on the results. 

* Tested wether specifying the nu-function improved the fit (reduced the AIC)

* Use Zero-inflated distributions to fit the model (available in the GAMLSS package)

#### GAM model CPUE - traps only
Statistical analysis of plot of trap catch rates. 
  
Because  traps and gillnets are very different & their deployment is confounded with depth (gillnets fish at surface, traps fishat bottom) it is not appropriate to combine both gear into a single model, they must be evaluated seperately. 

```{r GAM Traps only, echo=FALSE, warning=FALSE, message=FALSE}
cpue.st <- subset(catch.cpue, gear=="TB" | gear=="GN") %>% 
  dplyr::group_by(ss) %>% 
  dplyr::summarise(id=first(id), Fhrs=first(Fhrs), CPUEw=sum(CPUEw), CPUEn=sum(CPUEn), weight=sum(weight), number=sum(number), depth=first(depth), gear=first(gear), id=first(id), survey=first(survey), gear=first(gear), survey_m=first(survey_m))
cpue.st$survey_m <- as.factor(cpue.st$survey_m)

c22 <- subset(cpue.st, gear=="TB")

# Model of CPUEw with survey as random effect
mod22<-gamlss(CPUEw~depth+factor(id)+random(factor(survey)),family=ZAGA, data=c22) 

#model of CPUEw with survey as factor
mod33 <- gamlss(CPUEw~depth+factor(id)+factor(survey),family=ZAGA, data=c22) 

#model of CPUEn with survey as random factor
mod44 <-gamlss(CPUEn~depth+factor(id)+random(factor(survey)),family=ZAGA, data=c22) 

#model of CPUEn with survey as factor
mod55 <- gamlss(CPUEn~depth+factor(id)+factor(survey),family=ZAGA, data=c22) 

#Specifying nu-function for mod55 and mod33

mod333 <- gamlss(CPUEw~depth+factor(id)+factor(survey),family=ZAGA, nu.fo=~depth+factor(id)+factor(survey), data=c22) 

mod555 <- gamlss(CPUEn~depth+factor(id)+factor(survey),family=ZAGA, nu.fo=~depth+factor(id)+factor(survey), data=c22)


AIC(mod22, mod33, mod44, mod55, mod333, mod555)
#CPUEn:mod555 had the lowest AIC
#CPUEw: mod333 had the lowest AIC
summary(mod555) #CPUEn model
drop1(mod555)
plot(mod555)


summary(mod333) #CPUEw model
drop1(mod333)
plot(mod333)

```
**Trap CPUE GAM models**
The CPUEn models had better fit than the CPUEw models, and models with nu-function specified had better fit than those without.  

The overall bestfitting  model was mod555 of CPUEn with survey as an ordniary factor both area and survey were significant variable (when excluding the smoothing function), but not depth.  All survey and areas 1 & 5 were  significant factors in the model 

For the CPUEw model (mod333), only surveys was a significant varible, with only the Nov.2012 (intercept), and Nov.2013 survey being significant factors in the model.  

Conclusion: difference in trap CPUE by surveys was significantly different, but not so between areas. 


#### GAM model CPUE - gillnets only

```{r GAM gillnets only, echo=FALSE, warning=FALSE, message=FALSE}

c222 <- subset(cpue.st, gear=="GN")

# Model of CPUEw with survey as random effect
mod22<-gamlss(CPUEw~depth+factor(id)+random(factor(survey)),family=ZAGA, data=c222) 

#model of CPUEw with survey as factor
mod33 <- gamlss(CPUEw~depth+factor(id)+factor(survey),family=ZAGA, data=c222) 

#model of CPUEn with survey as random factor
mod44 <-gamlss(CPUEn~depth+factor(id)+random(factor(survey)),family=ZAGA, data=c222) 

#model of CPUEn with survey as factor
mod55 <- gamlss(CPUEn~depth+factor(id)+factor(survey),family=ZAGA, data=c222) 

#Specifying nu-function for mod55 and mod33

mod333 <- gamlss(CPUEw~depth+factor(id)+factor(survey),family=ZAGA, nu.fo=~depth+factor(id)+factor(survey), data=c222) 

mod555 <- gamlss(CPUEn~depth+factor(id)+factor(survey),family=ZAGA, nu.fo=~depth+factor(id)+factor(survey), data=c222)


AIC(mod22, mod33, mod44, mod55, mod333, mod555)
#CPUEn:mod555 had the lowest AIC
#CPUEw: mod333 had the lowest AIC
summary(mod333) #CPUEw model - best overall fit
drop1(mod333)
plot(mod333)

summary(mod555) #CPUEn model
drop1(mod555)
plot(mod555)


```
**Gillnet CPUE GAM models**
The gillnet GAM models had much lower AICs than the trap models, indicating a better overall model fit for gillnet data than for trap data.  
  
All CPUEw models had lower AIC than the CPUEn model (in contrast to the traps GAM models)  


The best fitting model was the CPUEw model with survey as an ordinary factor and the nu.function specified (mod333). Only area 3 & 6 were significant factors in the model, and in contrast to the traps surveys were not a significant factors. 

Conclusion: there were no significant differences in CPUE (w) for gillnets between surveys, and only significant differences involving areas 3 and 6. 



*** 

#### Bottom depth traps and gillnets
```{r freqplot trap and gillnet depth, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(cpue.st, aes(depth, colour = gear)) +  geom_freqpoly(binwidth = 5) +
  labs(title = "Bottom depth of trap and gill-net stations",
              subtitle = "i.e. fishing depth for traps, while gillnets were floating at the surface and fishing at 0-m depth") 

t.test(subset(cpue.st, gear=="TB")$depth, (subset(cpue.st, gear=="GN")$depth))

```

Depth was, however, not a significant factor in any of the four models developed for the trap CPUE, by weight or numbers. The depth distribution of bottom depth at the trap and gillnet stations (Gillnet fishing depth was at the surface, while trap fishing depth = bottom depth) were significantly different (t-test, p<0.05, df=67.9 -  see also frequncy plot above). Therefore, the models including both gear types led to a wider span and a more diverse distribution of the depths than for the traps-only data set. 

***

### Combine trap and gillnet duration + CPUE plots

```{r combined trap and gillnet plot, echo=FALSE, warning=FALSE, message=FALSE}

(duration.plot + ca.plot)/(duration.plot.gn | ca.plot2) + plot_annotation(tag_levels = 'A') &   theme(plot.tag = element_text(size = 10)) +theme_minimal()

ggsave("./figs/Fig3_traps_nets_4_panel.png")
ggsave("./figs/Fig3_traps_nets_4_panel.tiff")


```


*** 

### CPUE by gear type  - by family and trophic group (gearplot)
Only for gillnets and traps and stations with catches (excluding no-catch)

#### TRAP CATCHES by depth bins (0-30m and >30m)

Split CPUEw by depth ranges (0-30m and >30m) for the trap catches. Gillnets were all taken at the surface so belong in the upper depth range.
```{r gearplot by depth, echo=FALSE, warning=FALSE, message=FALSE}

c4 <- subset(catch.traits, TGShort!="NA" & (gear=="TB" | gear=="GN") & CPUEw>0)

############
# Depth range 0 - 30m

c41 <- subset(c4, (depth<=30 & gear=="TB"))
c42 <- subset(c4, (depth>30 & gear=="TB"))
c411 <- subset (c4, gear=="GN")
c41$Drng <- c("<30m")
c411$Drng <- c("<30m")
c42$Drng <- c(">30m")

c43 <- rbind(c41,c411, c42)


gear.plot.dr <- c43 %>%  dplyr::mutate(FamGroup2 = forcats::fct_relevel(FamGroup2,  "Other Spp", "Serranidae","Scombridae","Lutjanidae", "Lethrinidae", "Chirocentridae", "Carcharhinidae", "Carangidae",  "Acanthuridae")) %>%
 ggplot( aes(FamGroup2, fill=Drng)) + geom_bar(binwidt=0.2, aes( weight=CPUEw)) +theme_bw() + facet_grid(vars(gear), vars(survey_m), labeller=labeller(survey_m=c("Survey 1: Nov. 2012" = "Nov.2012",
      "Survey 2: May 2013" = "May 2013",
      "Survey 3: Nov. 2013" = "Nov.2013"), gear=c("GN" ="Gillnet", "TB" ="Trap"))) + coord_flip() +xlab("") + ylab("kg/hour fishing (cumulative)")  + guides(fill=guide_legend(title="Survey")) + theme(legend.position = "none") + guides(fill=FALSE) + scale_fill_manual(values=c("seagreen3", "dodgerblue4"))
gear.plot.dr
#ggsave("./figs/gearplot_species.tiff")


# by trophic group
gear.plot2.dr <- ggplot(c43, aes(TGShort, fill=Drng)) + geom_bar(binwidt=0.2, aes( weight=CPUEw)) +theme_bw() + facet_grid(vars(gear), vars(survey_m), labeller=labeller(survey_m=c("Survey 1: Nov. 2012" = "Nov.2012",
      "Survey 2: May 2013" = "May 2013",
      "Survey 3: Nov. 2013" = "Nov.2013"), gear=c("GN" ="Gillnet", "TB" ="Trap"))) + coord_flip() +xlab("") + ylab("kg/hour fishing (cumulative)")  + guides(fill=guide_legend(title="Survey")) + theme(legend.position = "none") + guides(fill=FALSE) + scale_fill_manual(values=c("seagreen3", "dodgerblue4"))
gear.plot2.dr
#ggsave("./figs/gearplot_trophic_group.tiff")


# Combine plots in patchwork 
pw2 <- gear.plot2.dr / gear.plot.dr + plot_layout(guides = "keep") + plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10)) & 
  theme(text = element_text(size=5)) +theme(axis.text.y = element_text(size=8))  +theme_minimal() + theme(panel.border = element_rect(linetype = "solid", fill = NA))
pw2  
ggsave("./figs/Fig 4 gearplot_comb.tiff")
ggsave("./figs/Fig 4 gearplot_comb.png")

```



### Multivariate analysis of catch composition
PCA analysis was carried out to explore the variability in the catch data (CPUE-by-weight) in relation to:
- survey
- gear
- area
- Trophic group


PCA was carried out on average catches (CPUE) per survey and area (needed to create a M*N matrix - cannot have uneven number of observations per group).  

#### PCA Gillnets and traps, no depth bins
```{r PCA gillnets trap CPUEw, echo=FALSE, warning=FALSE, message=FALSE}

# By trophic group 
# CPUEw
c5 <- subset(c4, CPUEw>0)[,c(1,16, 30, 24, 40)]
#c5 <- subset(c43, CPUEw>0)[,c(1,16, 30, 24, 39, 43)]
c5$index <- paste(c5$gear, c5$survey, c5$id, sep="_")

cc5 <- c5[,c(6,4,5)] %>%
#cc5 <- c5[,c(6,4,5,7)] %>%
  group_by(index, TGShort) %>%
  summarise(CPUEw=mean(CPUEw))%>%
  pivot_wider(names_from = TGShort, values_from = CPUEw, values_fill = list(CPUEw = 0)) 
cc5<-as.data.frame(cc5)
rownames(cc5) <- cc5$index

cc5$index <- NULL

df_pca <- prcomp(cc5)

plot(df_pca$x[,1], df_pca$x[,2])

df_out <- as.data.frame(df_pca$x)
df_out$gear <- sapply( strsplit(as.character(row.names(cc5)), "_"), "[[", 1 )
df_out$survey <- sapply( strsplit(as.character(row.names(cc5)), "_"), "[[", 2 )
df_out$area <- sapply( strsplit(as.character(row.names(cc5)), "_"), "[[", 3 )
#df_out$depth <- sapply( strsplit(as.character(row.names(cc5)), "_"), "[[", 4 )
head(df_out)

library(grid) 
library(gridExtra)

percentage <- round(df_pca$sdev / sum(df_pca$sdev) * 100, 2)
percentage <- paste( colnames(df_out), "(", paste( as.character(percentage), "%", ")", sep="") )

# score plot - by gear and survey
p<-ggplot(df_out,aes(x=PC1,y=PC2,color=gear, label=area, shape = factor(survey)))
p<-p +
  geom_point() + 
  #geom_text(size=3,  nudge_x = 0.15) + 
  xlab(percentage[1]) + 
  ylab(percentage[2]) + 
  scale_colour_discrete(name  ="Gear",
                            breaks=c("GN", "TB"),
                            labels=c("Gillnet", "Traps")) +
      scale_shape_discrete(name  ="Survey",
                           breaks=c("2012901", "2013002", "2013005"),
                           labels=c("Nov.2012", "May 2013", "Nov. 2013")) +  theme_bw()
  
#score plot by gear and area
p1<-ggplot(df_out,aes(x=PC1,y=PC2,color=gear, label=area, shape = factor(survey)))
p1<-p1 +
  geom_text(position=position_jitter(width=0.2,height=0.2)) + 
  #geom_text(size=3,  nudge_x = 0.15) + 
  xlab(percentage[1]) + 
  ylab(percentage[2]) + 
  scale_colour_discrete(name  ="Gear",
                            breaks=c("GN", "TB"),
                            labels=c("Gillnet", "Traps")) +
      scale_shape_discrete(name  ="Survey",
                           breaks=c("2012901", "2013002", "2013005"),
                           labels=c("Nov.2012", "May 2013", "Nov. 2013")) +  theme_bw() + theme(legend.position = "none")
p1


df_out_r <- as.data.frame(df_pca$rotation)
df_out_r$Trophic_Group <- row.names(df_out_r)

df_out_r

#loadings plot
p2<-ggplot(df_out_r,aes(x=PC1,y=PC2,label=Trophic_Group ))
p2<-p2+ geom_text(size=3,position=position_jitter(width=0.2,height=0.2)) +  expand_limits(x = c(-0.15, 1.12), y=c(-1.12, 0.12)) +
  xlab(percentage[1]) + 
  ylab(percentage[2])+ theme_bw() + theme(legend.position = "none")



pw3 <- (p +p2) / p1
pw3

#ggsave("./figs/PCA_catchcomp_weight.tiff", width=6, height=6)
#ggsave("./figs/PCA_catchcomp_weight.png", width=6, height=6)
```

#### PCA of TRAP catches split in depth categories (<30m and >30m)
(gillnets were only at the surface)

```{r PCA traps by depth, echo=FALSE, warning=FALSE, message=FALSE}
c5 <- subset(c43, (CPUEw>0 & gear=="TB"))[,c(1,16, 30, 24, 40, 43)]
c5$index <- paste(c5$gear, c5$survey, c5$id, c5$Drng, sep="_")


cc5 <- c5[,c(6,4,5,7)] %>%
  group_by(index, TGShort) %>%
  summarise(CPUEw=mean(CPUEw))%>%
  pivot_wider(names_from = TGShort, values_from = CPUEw, values_fill = list(CPUEw = 0)) 
cc5<-as.data.frame(cc5)
rownames(cc5) <- cc5$index

cc5$index <- NULL

df_pca <- prcomp(cc5)

plot(df_pca$x[,1], df_pca$x[,2])

df_out <- as.data.frame(df_pca$x)
#df_out$gear <- sapply( strsplit(as.character(row.names(cc5)), "_"), "[[", 1 )
df_out$survey <- sapply( strsplit(as.character(row.names(cc5)), "_"), "[[", 2 )
df_out$area <- sapply( strsplit(as.character(row.names(cc5)), "_"), "[[", 3 )
df_out$depth <- sapply( strsplit(as.character(row.names(cc5)), "_"), "[[", 4 )
head(df_out)

library(grid) 
library(gridExtra)

percentage <- round(df_pca$sdev / sum(df_pca$sdev) * 100, 2)
percentage <- paste( colnames(df_out), "(", paste( as.character(percentage), "%", ")", sep="") )

#score plot by gear and survey
p.d<-ggplot(df_out,aes(x=PC1,y=PC2,color=depth, label=area, shape = factor(survey)))
p.d<-p.d +
  geom_point() + 
  #geom_text(size=3,  nudge_x = 0.15) + 
  xlab(percentage[1]) + 
  ylab(percentage[2]) + 
  scale_colour_manual(name  ="Depth range",
                            breaks=c("<30m", ">30m"),
                            labels=c("Shallow (<30m)", "Deep (>30m)"), 
                            values=c("seagreen3", "dodgerblue4")) +
      scale_shape_discrete(guide=FALSE, name  ="Survey",
                           breaks=c("2012901", "2013002", "2013005"),
                           labels=c("Nov.2012", "May 2013", "Nov. 2013")) +
  theme_bw()
  

#score plot by gear and depth range
p1.d<-ggplot(df_out,aes(x=PC1,y=PC2,color=depth, label=area, shape = factor(survey)))
p1.d<-p1.d +
  geom_text(position=position_jitter(width=0.02,height=0.08)) + 
  #geom_text(size=3,  nudge_x = 0.15) + 
  xlab(percentage[1]) + 
  ylab(percentage[2]) + 
  scale_colour_manual(name  ="Depth range",
                            breaks=c("<30m", ">30m"),
                            labels=c("Shallow (<30m)", "Deep (>30m)"), 
                            values=c("seagreen3", "dodgerblue4")) +
      scale_shape_discrete(guide=FALSE, name  ="Survey",
                           breaks=c("2012901", "2013002", "2013005"),
                           labels=c("Nov.2012", "May 2013", "Nov. 2013")) +  theme_bw() + theme(legend.position = "none")
p1.d

df_out_r <- as.data.frame(df_pca$rotation)
df_out_r$Trophic_Group <- row.names(df_out_r)

df_out_r

p2.d<-ggplot(df_out_r,aes(x=PC1,y=PC2,label=Trophic_Group ))
p2.d<-p2.d+ geom_text(size=3,position=position_jitter(width=0.2,height=0.25)) +  expand_limits(x = c(-0.17, 1.12), y=c(-1.12, 0.12)) +
  xlab(percentage[1]) + 
  ylab(percentage[2])+ theme_bw() + theme(legend.position = "none")

pw3.d <- (p.d+p2.d) / p1.d
pw3.d

```



##### Combine PCA plots
```{r gearplot pca, echo=FALSE, warning=FALSE, message=FALSE}
#pw2 | pw3 

(p + p.d) / (p1 +p1.d) / (p2 + p2.d) + plot_layout(guides = 'collect') + plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 9)) 
 


#gear.plot2| pw3 

ggsave("./figs/Fig5_PCA.tiff", width=12, height=10)
ggsave("./figs/Fig5_PCA.png", width=12, height=10)
```
In Figure 4 panels A and B clearly show how the catch composition differed between traps and gillnets, both in relation to trophic groups and main fish families caught.  Carinvores were caught in similar quantities by both the traps and gillnets, coralivores only in traps (but in very low quantities) invertivores mainly in traps, planktivores and herbivores mainly in gillnets.  

Catch composition by families (Figure 4 panel A and B) showed clear differences in the species selectivity of  the gear types. 

Catch composition of traps showed some differences by depth range (<30m or > 30m), for the plantivores trophic group and Acunthuridae and Carangidae families that were only caught at shallow depths, while no trophic groups or families were uniquely caught at depths greater than 30m. 


PCA of catch weight and numbers (CPUEw and CPUEw) by trophic group  gave similar results and showed how trap and gillnet catches differed. The following analysis therefore focused solely on PCA by CPUEw.  

In the PCA analysis of both gillnet and traps (Figure 4, panel C and D). The first two principal components explained 88% of the variation. 
A large gillnet catch of Planktivores (56.69 kg gillnet catch of Naso hexacanthus in area 2 during the Nov. 2013 survey) drove the variability along PC1. PC2 was driven by high  catches of carnivores  during the November 2013 survey d. Catches of coralivores, invertivores and herbivores contributed little to the variability in either of the two first PCs (but these were caught in low quantities in just a few surveys and areas).  

The PCA analysis of the trap catches (Figure 4, panel E and F) the first two PCs explained 67.9% of the variability with PC1 and PC2 contributing fairly equally (35.9% and 32% respectively). Catches of carnivores and invertivores drove the variability along PC1 while a 10.14kg catch of planktivores (Naso hexacanthus) in area 5 in May 2013 in shallow waters  drove the variability along PC2. Apart from the May.2013 point no areas showed particular differences between deep and shallow areas. 
  
Trap catches were dominated by herbivores, gillnet catches by carnivores, while gillnet catches in area 2 during the Nov. 2013 survey had very high catches of planktivores compared to other gears, surveys and areas. Trap catches did not show marked difference by depth category, except for Area 5 during the May 2013 survey where there were (relatively) large catches of planktivores. 


However, this is still descriptive. Need to evaluate the differences statistically. Will use GAM model.   
(This also allows to use data on station level, not just average catches per area/survey needed to construct a matrix necessary for PCA)


***

#### Catch composition - GAM Models CPUEw 
Expand on the PCA analysis and descriptive plots using GAM models.  Compartive, evaluating the same variables as in the PCA analysis:
- Survey
- Area
- Depth Category (shallow / deep)
- Trophic group

Separate models for Gillnets and Traps,adding depth category as factor variable.  


##### Trap catch comp GAMs
```{r  GAM CPUEw traps by all factor, echo=FALSE, warning=FALSE, message=FALSE}
## TRAPS
#find best-fitting distribution
ma <- gamlss(CPUEw~factor(Drng)+factor(id)+random(factor(survey))+factor(TGShort),family=NO, data=na.omit(subset(c43, gear=="TB" & CPUEw>0))) 

t1 <- chooseDist(ma, type="realplus")
getOrder(t1,3)
#cmod1<-update(ma, family=names(getOrder(t1,3)[1]))

# traps only, CPUEw, survey as random factor
mod7 <- gamlss(CPUEw~factor(Drng)+factor(id)+random(factor(survey))+factor(TGShort),family=names(getOrder(t1,3)[1]), data=na.omit(subset(c43, gear=="TB" & CPUEw>0))) 
#summary(mod7)

# traps only, CPUE by numbers, survey as ordinary factor
mod8 <- gamlss(CPUEw~factor(Drng)+factor(id)+factor(survey)+factor(TGShort),family=names(getOrder(t1,3)[1]), data=na.omit(subset(c43, gear=="TB" & CPUEw>0)))
#summary(mod8)

#without surveys
 mod10 <- gamlss(CPUEw~factor(Drng)+factor(id)+factor(TGShort),family=names(getOrder(t1,3)[1]), data=na.omit(subset(c43, gear=="TB" & CPUEw>0))) 
 
 # without surveys and depth
 mod12 <- gamlss(CPUEw~factor(id)+factor(survey)+factor(TGShort),family=names(getOrder(t1,3)[1]), data=na.omit(subset(c43, gear=="TB" & CPUEw>0)))
 

AIC(mod7, mod8, mod10, mod12)
#mod10 had lowest AIC
summary(mod10)
drop1(mod10)
```

 
##### Gillnet catch comp GAMs
Excluding depth as a factor
```{r gillnet catch comp GAM, echo=FALSE, warning=FALSE, message=FALSE}
#find best-fitting distribution
ma <- gamlss(CPUEw~factor(id)+random(factor(survey))+factor(TGShort),family=NO, data=na.omit(subset(c43, gear=="GN" & CPUEw>0))) 

t1 <- chooseDist(ma, type="realplus")
getOrder(t1,3)
#cmod1<-update(ma, family=names(getOrder(t1,3)[1]))

# traps only, CPUEw, survey as random factor
mod7a <- gamlss(CPUEw~factor(id)+random(factor(survey))+factor(TGShort),family=names(getOrder(t1,3)[1]), data=na.omit(subset(c43, gear=="GN" & CPUEw>0))) 
#summary(mod7)

# traps only, CPUE by numbers, survey as ordinary factor
mod8a <- gamlss(CPUEw~factor(id)+factor(survey)+factor(TGShort),family=names(getOrder(t1,3)[1]), data=na.omit(subset(c43, gear=="GN" & CPUEw>0)))
#summary(mod8)

#without surveys
 mod10a <- gamlss(CPUEw~factor(id)+factor(TGShort),family=names(getOrder(t1,3)[1]), data=na.omit(subset(c43, gear=="GN" & CPUEw>0))) 
 
 # without surveys and depth
 mod12a <- gamlss(CPUEw~factor(survey)+factor(TGShort),family=names(getOrder(t1,3)[1]), data=na.omit(subset(c43, gear=="GN" & CPUEw>0)))
 

AIC(mod7a, mod8a, mod10a, mod12a)
#mod10 had lowest AIC
summary(mod7a)
drop1(mod7a)


```

####Hyperstability trap catches - max number of fish per trap
```{r trap CPUE hyperstability, echo=FALSE, warning=FALSE, message=FALSE}

c5 <- subset(c4, gear=="TB" & CPUEn>0)
#hist(c5$number)

#sort(c5$number)

ggplot(c5, aes(sort(number))) + geom_bar()

```
Asymptotic histogram - no indication of reaching a plateau, only low probability of catching many fish. 

### Test differences in CPUE By Species in trap catches, by range of traits 
Use GAM model and evaluate how CPUE by weight (CPUEw) was affected by the traits :

- trophic group
- trophic level (linked to trophic group)
- Place in water column
- Diel activity
- Habitat 
- Gregariousness
- MaxLength

#### GAM CPUEw - all traits
(excluding depth, area and survey)
```{r GAM CPUEw by all traits, echo=FALSE, warning=FALSE, message=FALSE}
ct <- subset(c4, gear=="TB" & CPUEn>0)
#ct <- ct[,c(1,22,23,25,28,29,30,31,32,33,34,35,37)]

#find best-fitting distribution
#ma <- gamlss(CPUEn~depth+factor(id)+factor(survey)+TrophicLevel+factor(TGShort)+factor(WaterCol)+factor(DielActivity)+factor(Habitat)+factor(Gregariousness)+MaxLength,family=NO, data=na.omit(ct)) 
ma <- gamlss(CPUEn~TrophicLevel+factor(TGShort)+factor(WaterCol)+factor(DielActivity)+factor(Habitat)+factor(Gregariousness)+MaxLength,family=NO, data=na.omit(ct)) 

t1 <- chooseDist(ma, type="realplus")
getOrder(t1,3)
#cmod1<-update(ma, family=names(getOrder(t1,3)[1]))

# traps only, CPUE by numbers, survey as ordinary factor
ma <- gamlss(CPUEn~TrophicLevel+factor(TGShort)+factor(WaterCol)+factor(DielActivity)+factor(Habitat)+factor(Gregariousness)+MaxLength,family=names(getOrder(t1,3)[1]), data=na.omit(ct)) 

summary(ma)
drop1(ma)
plot(ma)

```
Evaluting catches of fish in relation to their traits. Should be evaluated by numbers, for single large fish not to dominate the results. The model based on CPUE by numbers (mod12, where surveys, area and depth are excluded) had  lower AIC than models with numbers and fishing hours as a factor.  Also, including survey as an ordinary factor led to lower AIC than including it as a random factor. 

Model (BCPOe Box-Cox Power Exponential-orig. distribution, AIC: -1384.861, df:17) results show that numbers of fish caught (CPUE by numbers) was significantly affected by:  

* all traits, except MaxLength and WaterColumn (demersal)

 

### CPUE by trophic group by area & by survey (year) 

Presents the CPUE pr hr, for each management area and survey, standardized by the number of traps set in each area X survey combination. 

Trap catches were dominated by carnivores, occuring in all area X survey combinations, followed by invertivores who also occurred everywhere, albeight in lower densities than the carnivores. 

Cathces of planktivores and invertivores varied substantially between areas and surveys, in contrast to carnivores who were caught in all surveys and areas. This explains why invertivores and planktivores were signifant factors on the catches (GAM model of CPUEn with traits), while carnivores were not. 

*Area 2, 5 and 6 seem to have the highest diversity of trophic groups, but this has to be calculated as functional diversity, using the FD package (or similar). *

```{r trophic plot, echo=FALSE, warning=FALSE, message=FALSE}

trophic.plot <- ggplot(subset(catch.traits, TGShort!="NA" & gear=="TB"), aes(TGShort, CPUEw/NArea))

trophic.plot <- trophic.plot + geom_col(aes(fill=TGShort)) + facet_grid(vars(id), vars(survey_m)) + scale_fill_manual(name="Trophic group",values=qualitative_hcl(n = 5, h = c(0, 295), c = 80, l = 60, register = ) ) + labs(x=" ", y="CPUE (kg/hour fishing) / (number of traps pr area)") + theme_bw() 
trophic.plot
ggsave("./figs/S2_trophic_group_cpue.png")

tiff(file="./figs/S3_trophic_group_cpue.tiff", width=3000, height=3000, res=400, pointsize=10, compression=c("none"))
trophic.plot
dev.off()
```


### Trophic group CPUE by depth
A modified version of Figure 6 in Feb 2019 MS. 

Basically a more complex version of the previous plot. Adds the depth dimension to the plot, and the surveys as colours. 

Shows uniform catch rates of carnivores and invertivores from 0-50m. 

All surveys seem to overlap, so no difference in trophich group depth dependent catch rates between the surveys.

```{r trophic depth plot, echo=FALSE, warning=FALSE, message=FALSE}

depth.plot <- ggplot(subset(catch.traits, gear=="TB" & TGShort!="NA"), aes(depth, CPUEw)) 

depth.plot <- depth.plot + geom_point(aes(colour=survey_m))  + facet_grid(vars(id),vars(TGShort)) +  scale_colour_manual(name="Survey",values=qualitative_hcl(n = 3, h = c(0, 295), c = 80, l = 60, register = ) ) + labs(x="Depth (m)", y="CPUE (kg(hrs)")  +theme_bw()  

depth.plot

tiff(file="./figs/trophic_group_cpue_depth.tiff", width=3000, height=3000, res=400, pointsize=10, compression=c("none"))
depth.plot
dev.off()
```


### Trophic level in Trap catches
```{r trophic level, echo=FALSE, warning=FALSE, message=FALSE}

TMean <- subset(catch.traits, gear=="TB" & TrophicLevel!="NA") %>% group_by(id, survey_m) %>% summarise(TM=mean(na.omit(TrophicLevel)))
CMean <- subset(catch.traits, gear=="TB" & TrophicLevel!="NA") %>% group_by(id, survey_m) %>% summarise(CM=mean(CPUEw))
TMean$CM <- CMean$CM

t0.plot <- ggplot(subset(catch.traits, gear=="TB" & TrophicLevel!="NA"), aes(TrophicLevel, CPUEw)) 

t0.plot <- t0.plot + geom_point(aes(colour=TGShort)) + geom_vline(data = TMean, aes(xintercept = TM)) + facet_grid(vars(id),vars(survey_m)) +  scale_colour_manual(name="Trophic group",values=qualitative_hcl(n = 5, h = c(0, 295), c = 80, l = 60, register = ) ) + labs(x="Trophic level", y="CPUE (kg(hrs)")  +theme_bw()  
t0.plot

tiff(file="./figs/trophic_level.tiff", width=3000, height=3000, res=400, pointsize=10, compression=c("none"))
t0.plot
dev.off()
```

#### GAM model of trophic level of catches



```{r GAM trophic level, echo=FALSE, warning=FALSE, message=FALSE}
ct <- subset(catch.traits, gear=="TB" & TrophicLevel!="NA")
#ct <- ct[,c(1,25,28,30)]

tl1 <- gamlss(TrophicLevel~depth+factor(id)+factor(survey), family=NO, data=na.omit(ct))

t11 <- chooseDist(tl1, type="realplus")
getOrder(t11,3)

names(getOrder(t1,3)[1])

tl1 <- gamlss(TrophicLevel~depth+factor(id)+factor(survey), family=names(getOrder(t1,3)[1]), data=na.omit(ct))

# tested other models, nu-fu does not improve AIC


summary(tl1)
drop1(tl1)
#plot(tl1)

```
The GAM Model (Box-Cox Power Exponential, AIC: 144.0248, df: 12) of trophic level by depth, area and surveys only showed area 3 and the intercept to be significant factors affecting the Trophic level of catches meaning that trophic level was similar in all areas except area 3 (and 1)


### Maxlength in catches
```{r maxlength, echo=FALSE, warning=FALSE, message=FALSE}

t0.plot <- ggplot(subset(catch.traits, gear=="TB" & TrophicLevel!="NA"), aes(MaxLength, CPUEn)) 

t0.plot <- t0.plot + geom_point(aes(colour=TGShort)) + facet_grid(vars(id),vars(survey_m)) +  scale_colour_manual(name="Trophic group",values=qualitative_hcl(n = 5, h = c(0, 295), c = 80, l = 60, register = ) ) + labs(x="Max lenght(cm)", y="CPUE (no. fish/hour)")  +theme_bw()  
t0.plot
ggsave("./figs/S6_max_length.png")

tiff(file="./figs/S7_max_length.tiff", width=3000, height=3000, res=400, pointsize=10, compression=c("none"))
t0.plot
dev.off()


```


### Place in water column
```{r place in water column, echo=FALSE, warning=FALSE, message=FALSE}

catch.traits$WaterCol[catch.traits$WaterCol=="Demersal"] <- "demersal"
tl.plot <- ggplot(subset(catch.traits, WaterCol!="NA" & gear=="TB"), aes(WaterCol, CPUEw/NArea))

tl.plot <- tl.plot + geom_col(aes(fill=WaterCol)) + facet_grid(vars(id), vars(survey_m)) + scale_fill_manual(name="Place in water column",values=qualitative_hcl(n = 6, h = c(0, 295), c = 80, l = 60, register = ) ) + labs(x=" ", y="CPUE (kg/hour fishing) / number of traps pr area") + theme_bw() + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
tl.plot
ggsave("./figs/S3_water_col.png")

tiff(file="./figs/S4_water_col.tiff", width=3000, height=3000, res=400, pointsize=10, compression=c("none"))
tl.plot
dev.off()
```

### Diel activity
```{r diel activity, echo=FALSE, warning=FALSE, message=FALSE}

t2.plot <- ggplot(subset(catch.traits, DielActivity!="#N/A" & gear=="TB"), aes(DielActivity, CPUEw/NArea))

t2.plot <- t2.plot + geom_col(aes(fill=DielActivity)) + facet_grid(vars(id), vars(survey_m)) + scale_fill_manual(name="Diel activity",values=qualitative_hcl(n = 6, h = c(0, 295), c = 80, l = 60, register = ) ) + labs(x=" ", y="CPUE (kg/hour fishing) / number of traps pr area") + theme_bw() + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
t2.plot
ggsave("./figs/S4_diel_act.png")

tiff(file="./figs/S5_diel_act.tiff", width=3000, height=3000, res=400, pointsize=10, compression=c("none"))
t2.plot
dev.off()
```


### Gregariousness
```{r Gregariousness, echo=FALSE, warning=FALSE, message=FALSE}

t3.plot <- ggplot(subset(catch.traits, Gregariousness!="#N/A" & gear=="TB"), aes(Gregariousness, CPUEw/NArea))

t3.plot <- t3.plot + geom_col(aes(fill=Gregariousness)) + facet_grid(vars(id), vars(survey_m)) + scale_fill_manual(name="Gregariousness",values=qualitative_hcl(n = 6, h = c(0, 295), c = 80, l = 60, register = ) ) + labs(x=" ", y="CPUE (kg/hour fishing) / number of traps pr area") + theme_bw() + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
t3.plot

ggsave("./figs/S5_gregariousness.png")

tiff(file="./figs/S6_gregariousness.tiff", width=3000, height=3000, res=400, pointsize=10, compression=c("none"))
t3.plot
dev.off()
```




***
## Biodiversity (species densities, trophic diversity etc)

### Species accumulation curves
```{r species accumulation, echo=FALSE, warning=FALSE, message=FALSE}

c41b <- subset(catch.unmod, (depth<=30 & gear=="TB"))
c42b <- subset(catch.unmod, (depth>30 & gear=="TB"))
c411b <- subset (catch.unmod, gear=="GN")
c41b$Drng <- c("<30m")
c411b$Drng <- c("<30m")
c42b$Drng <- c(">30m")

c43b <- rbind(c41b,c411b, c42b)

## TRAPS


### Shallow Traps (<30m)
catch2 <- subset(c43b, (gear=="TB" & species !="NOCATCH" & species != "FISJU09" & Drng=="<30m"))
catch2$Fh_int <- as.integer(catch2$Fhrs)

sp.acc.all <- data.frame(hrs=(min(catch2$Fh_int):max(catch2$Fh_int)), All=0, Area1=0, Area2=0, Area3=0, Area4=0, Area5=0, Area6=0, Area7=0)

for(i in 1:dim(sp.acc.all)[1]) {
  sp.acc.all[i,2] <- subset(catch2, Fh_int<=sp.acc.all[i,1]) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,3] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==1)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,4] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==2)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,5] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==3)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,6] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==4)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,7] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==5)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,8] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==6)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,9] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==7)) %>% dplyr::summarise(n_distinct(species))
}

sp.acc.all.long <- tidyr::gather(sp.acc.all, "area", "n", 2:9)

sp.traps.s <- ggplot(sp.acc.all.long, aes(x=hrs, y=n, group=area)) + geom_step(size=2, aes(colour=area)) + theme_classic() + scale_colour_brewer(palette = "Dark2") + xlab("Hours fishing") + ylab("Number of species") + theme(legend.position = "none") + ggtitle("A: Traps (<30m)")

shallow.acc <- sp.acc.all.long %>% group_by(area) %>% dplyr::summarise(spc=max(n))


### Deep Traps (>30m)
catch2 <- subset(c43b, (gear=="TB" & species !="NOCATCH" & species != "FISJU09" & Drng==">30m"))
catch2$Fh_int <- as.integer(catch2$Fhrs)

sp.acc.all <- data.frame(hrs=(min(catch2$Fh_int):max(catch2$Fh_int)), All=0, Area1=0, Area2=0, Area3=0, Area4=0, Area5=0, Area6=0, Area7=0)

for(i in 1:dim(sp.acc.all)[1]) {
  sp.acc.all[i,2] <- subset(catch2, Fh_int<=sp.acc.all[i,1]) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,3] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==1)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,4] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==2)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,5] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==3)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,6] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==4)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,7] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==5)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,8] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==6)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,9] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==7)) %>% dplyr::summarise(n_distinct(species))
}

sp.acc.all.long <- tidyr::gather(sp.acc.all, "area", "n", 2:9)

sp.traps.d <- ggplot(sp.acc.all.long, aes(x=hrs, y=n, group=area)) + geom_step(size=2, aes(colour=area)) + theme_classic() + scale_colour_brewer(palette = "Dark2") + xlab("Hours fishing") + ylab("Number of species") + theme(legend.position = "none") + ggtitle("B: Traps (>30m")

deep.acc <- sp.acc.all.long %>% group_by(area) %>% dplyr::summarise(spc=max(n))



## GILLNETS

catch2 <- subset(c43b, gear=="GN" & species !="NOCATCH" & species != "FISJU09")
catch2$Fh_int <- as.integer(catch2$Fhrs)

sp.acc.all <- data.frame(hrs=(min(catch2$Fh_int):max(catch2$Fh_int)), All=0, Area1=0, Area2=0, Area3=0, Area4=0, Area5=0, Area6=0, Area7=0)

for(i in 1:dim(sp.acc.all)[1]) {
  sp.acc.all[i,2] <- subset(catch2, Fh_int<=sp.acc.all[i,1]) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,3] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==1)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,4] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==2)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,5] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==3)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,6] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==4)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,7] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==5)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,8] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==6)) %>% dplyr::summarise(n_distinct(species))
  sp.acc.all[i,9] <- subset(catch2, (Fh_int<=sp.acc.all[i,1] & id==7)) %>% dplyr::summarise(n_distinct(species))
}

sp.acc.all.long <- tidyr::gather(sp.acc.all, "area", "n", 2:9)

sp.nets <- ggplot(sp.acc.all.long, aes(x=hrs, y=n, group=area)) + geom_step(size=2, aes(colour=area)) + theme_classic() + scale_colour_brewer(palette = "Dark2") + xlab("Hours fishing") + ylab("Number of species") + ggtitle("C: Gillnets")

gill.acc <- sp.acc.all.long %>% group_by(area) %>% dplyr::summarise(spc=max(n))

cbind(shallow.acc, deep.acc$spc, gill.acc$spc)

sp.comb <- sp.traps.s / sp.traps.d / sp.nets
sp.comb

ggsave("./figs/Fig 6 species_accumulation.png", height = 9, width = 6)
ggsave("./figs/Fig 6 species_accumulation.tiff", height = 9, width = 6)
```



*** 

### Functional diversity
UPDATE March 2021. Including both traps and gillnet data in analysis.  

We did not split the FD analysis by depth (<> 30m) as this would bias the shallow bin because that would include both gillnets and traps while the deep bin only included trap catches. 

We used the same approach as Stuart-Smith et al. to exclude very few observations of a single species and excluded all species observations occuring in 2 or fewer stations (traps). 

To avoid dbFD crashing 'm' was set in the 'dbFD' call (the number of PoCA  axes kept as traits to do the FRic analysis). Several levels of 'm' were tested, and this could probably be tested further (to higher levels of 'm')

##### Traps & Gillnets,  Results for n_spec > 2 m=10
Results based on the output from the dbFD analysis:

Overall Fric (R2)=  0.533 (compared to 0.62 for traps only data)


```{r functional diversity, echo=FALSE, warning=FALSE, message=FALSE}

#ss_list <- levels(catch1$survey)
#sp_occur <- subset(catch1, CPUEn>0 & gear=="TB")  %>% dplyr::group_by(Sci_name) %>% dplyr::summarise(n())
sp_occur <- subset(catch1, CPUEn>0 & gear=="TB" | gear=="GN")  %>% dplyr::group_by(Sci_name) %>% dplyr::summarise(n())
colnames(sp_occur) <- c("Sci_name", "n_spec")

# wrangle traits into a traitstable that can be used in FD
missing <- setdiff( x=traits$x1, y=subset(sp_occur, n_spec>2)$Sci_name)
t5 <- traits[!traits$x1 %in% missing,]
#t5 <- traits[traits$x1 %in% subset(sp_occur, n_spec>2)$Sci_name,]
tr <- t5[,c(4:10)]
rownames(tr) <- t5$x1
tr[(tr=="#N/A")] <- c(NA)
tr <- droplevels(tr)
tr$diel<-as.integer(as.factor(tr$Diel.Activity))
tr$diel[(tr$diel==2)] <- c(0)
tr <- tr[,c(1,2,3,4,6,7,8)]
#convert character variables to factors
tr$Trophic.group <- as.factor(tr$Trophic.group)
tr$Water.column <- as.factor(tr$Water.column)
tr$Habitat <- as.factor(tr$Habitat)
tr$Gregariousness <- as.factor(tr$Gregariousness)



# wrangle 'catch1' to aggregate CPUE pr area (row) pr species (columns)
missing <- setdiff( x=catch1$Sci_name, y=rownames(tr))
c5 <- catch1[!catch1$Sci_name %in% missing,]
#c_melt <- melt(c5[,c(28,11,22)], id=c("id", "Sci_name")) # weight by CPUEw
#c_melt <- melt(c5[,c(28,11,23)], id=c("id", "Sci_name")) # weight by CPUEn
c_melt <- melt(c5[,c(30,6,25)], id=c("id", "Sci_name")) # weight by CPUEn

#convert abundance to integer for better calculation performance
med_int1000 <- function(x) as.integer(median(x)*1000)
abund <- dcast(c_melt, id ~ Sci_name, value.var="value", fun=med_int1000) 
# to calculate with raw abundance data, use line below:
#abund <- dcast(c_melt, id ~ Sci_name, value.var="value", fun=median)
abund[is.na(abund)] <- 0
abund <- abund[,c(2:length(abund))]

#calculate FD using dbFD function, m set to 10 to avoid crach (int overload in QHULL)
FuncDiv <- dbFD(tr, abund, corr = c("lingoes"), m=10)

FDt <- as.data.frame(rbind(FuncDiv$FEve, FuncDiv$FDiv, FuncDiv$FDis, FuncDiv$RaoQ))
rownames(FDt) <- c("FEve", "FDiv", "FDis", "RaoQ")
colnames(FDt) <- c("A1", "A2", "A3", "A4", "A5", "A6", "A7")

FDt <- as.data.frame(t(round(FDt, digits=3)))

# Nice-looking table using 'kable', highlighting highest values in red
FDt %>%
  dplyr::mutate(
    Area = cell_spec(row.names(.), color = "black"),
    FEve = cell_spec(FEve, "html", color = ifelse(FEve >= max(FEve), "red", "black")),
    FDiv = cell_spec(FDiv, "html", color = ifelse(FDiv >= max(FDiv), "red", "black")),
    FDis = cell_spec(FDis, "html", color = ifelse(FDis >= max(FDis), "red", "black")),
    RaoQ = cell_spec(RaoQ, "html", color = ifelse(RaoQ >= max(RaoQ), "red", "black"))
  ) %>%
  dplyr::select(Area, FEve, FDiv, FDis, RaoQ) %>%
  kable(format = "html", escape = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

#### Rerunning the FD model excluding one trait at a time
Need to reduce 'm' to 9 to avoid crash:

*Zero distance(s)Error in convhulln(tr.FRic, "FA") :* 
  *Received error code 2 from qhull. Qhull error:*
    *QH6114 qhull precision error: initial simplex is not convex. Distance=1.4e-16* 
    
Removing traits only improved the R2 by max 9.6% (by removing Maxlength). Therefore decided to keep all the traits in the model for Traps and Gillnets
  
```{r FD model excluding traits, echo=FALSE, warning=FALSE, message=FALSE}
# create data frame with original RaoQ and FRic
Raoq <- as.data.frame(c(FuncDiv$RaoQ, FuncDiv$qual.FRic))
colnames(Raoq) <- c("All")
rownames(Raoq) <- c("A1", "A2", "A3", "A4", "A5", "A6", "A7", "R2")


# select tr, but excluding one variable at a time
#tr[,c(1,!2,3,4,5,6,7)]
for (i in 1:7) { 
  FDout <- dbFD(tr[,c(setdiff(1:7, i))], abund, corr = c("lingoes"), m=10)
  raoq1 <- as.data.frame(c(FDout$RaoQ, FDout$qual.FRic))
  colnames(raoq1) <- colnames(tr)[i]
  Raoq <- cbind(Raoq, raoq1)
}

#r2 for all models
formattable(round(Raoq[8,], digits=3))

#R2 per area and model
formattable(round(Raoq[c(1:7),], digits=3), list(    MaxLength = color_tile("transparent", "lightpink"),     Trophic.Level = color_tile("transparent", "lightpink"),     Trophic.group = color_tile("transparent", "lightpink"),    Water.column = color_tile("transparent", "lightpink"), Habitat = color_tile("transparent", "lightpink"),    Gregariousness = color_tile("transparent", "lightpink"), diel = color_tile("transparent", "lightpink"), ML_TL= color_tile("transparent", "lightpink")))

write.csv2(Raoq, "./figs/Raoq_table.csv")

```

#### Trap only - RaosQ
For traps excluding traits had an even less effect than for the traps & gillnet analysis. The only removal of trait that increased R2 was for trophic level, but the increase was only 0.7%. Therefore the full model including all traits was used in the manuscript. 
```{r trap only FD, echo=FALSE, warning=FALSE, message=FALSE}
sp_occur <- subset(catch1, CPUEn>0 & gear=="TB")  %>% dplyr::group_by(Sci_name) %>% dplyr::summarise(n())
colnames(sp_occur) <- c("Sci_name", "n_spec")

# wrangle traits into a traitstable that can be used in FD
missing <- setdiff( x=traits$x1, y=subset(sp_occur, n_spec>2)$Sci_name)
t5 <- traits[!traits$x1 %in% missing,]
tr <- t5[,c(4:10)]
rownames(tr) <- t5$x1
tr[(tr=="#N/A")] <- c(NA)
tr <- droplevels(tr)
tr$diel<-as.integer(as.factor(tr$Diel.Activity))
tr$diel[(tr$diel==2)] <- c(0)
tr <- tr[,c(1,2,3,4,6,7,8)]
#convert character variables to factors
tr$Trophic.group <- as.factor(tr$Trophic.group)
tr$Water.column <- as.factor(tr$Water.column)
tr$Habitat <- as.factor(tr$Habitat)
tr$Gregariousness <- as.factor(tr$Gregariousness)

# wrangle 'catch1' to aggregate CPUE pr area (row) pr species (columns)
missing <- setdiff( x=catch1$Sci_name, y=rownames(tr))
c5 <- catch1[!catch1$Sci_name %in% missing,]
c_melt <- melt(c5[,c(30,6,25)], id=c("id", "Sci_name")) # weight by CPUEn

#convert abundance to integer for better calculation performance
med_int1000 <- function(x) as.integer(median(x)*1000)
abund <- dcast(c_melt, id ~ Sci_name, value.var="value", fun=med_int1000) 
# to calculate with raw abundance data, use line below:
#abund <- dcast(c_melt, id ~ Sci_name, value.var="value", fun=median)
abund[is.na(abund)] <- 0
abund <- abund[,c(2:length(abund))]

#calculate FD using dbFD function, m set to 10 to avoid crach (int overload in QHULL)
FuncDiv <- dbFD(tr, abund, corr = c("lingoes"), m=10)

FDt <- as.data.frame(rbind(FuncDiv$FEve, FuncDiv$FDiv, FuncDiv$FDis, FuncDiv$RaoQ))
rownames(FDt) <- c("FEve", "FDiv", "FDis", "RaoQ")
colnames(FDt) <- c("A1", "A2", "A3", "A4", "A5", "A6", "A7")

FDt <- as.data.frame(t(round(FDt, digits=3)))

# Nice-looking table using 'kable', highlighting highest values in red
FDt %>%
  dplyr::mutate(
    Area = cell_spec(row.names(.), color = "black"),
    FEve = cell_spec(FEve, "html", color = ifelse(FEve >= max(FEve), "red", "black")),
    FDiv = cell_spec(FDiv, "html", color = ifelse(FDiv >= max(FDiv), "red", "black")),
    FDis = cell_spec(FDis, "html", color = ifelse(FDis >= max(FDis), "red", "black")),
    RaoQ = cell_spec(RaoQ, "html", color = ifelse(RaoQ >= max(RaoQ), "red", "black"))
  ) %>%
  dplyr::select(Area, FEve, FDiv, FDis, RaoQ) %>%
  kable(format = "html", escape = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

#### Test excluding one trait at a time
# create data frame with original RaoQ and FRic
Raoq2 <- as.data.frame(c(FuncDiv$RaoQ, FuncDiv$qual.FRic))
colnames(Raoq2) <- c("All")
rownames(Raoq2) <- c("A1", "A2", "A3", "A4", "A5", "A6", "A7", "R2")


# select tr, but excluding one variable at a time
#tr[,c(1,!2,3,4,5,6,7)]
for (i in 1:7) { 
  FDout <- dbFD(tr[,c(setdiff(1:7, i))], abund, corr = c("lingoes"), m=10)
  raoq1 <- as.data.frame(c(FDout$RaoQ, FDout$qual.FRic))
  colnames(raoq1) <- colnames(tr)[i]
  Raoq2 <- cbind(Raoq2, raoq1)
}

#r2 for all models
formattable(round(Raoq2[8,], digits=3))

#R2 per area and model
formattable(round(Raoq2[c(1:7),], digits=3), list(    MaxLength = color_tile("transparent", "lightpink"),     Trophic.Level = color_tile("transparent", "lightpink"),     Trophic.group = color_tile("transparent", "lightpink"),    Water.column = color_tile("transparent", "lightpink"), Habitat = color_tile("transparent", "lightpink"),    Gregariousness = color_tile("transparent", "lightpink"), diel = color_tile("transparent", "lightpink"), ML_TL= color_tile("transparent", "lightpink")))
```


***
## Results table
This table combines CPUE, traits pr gear type, species density and RaoQ into one results-table. 

```{r results table, echo=FALSE, message=FALSE}

st2 <- catch %>% dplyr::group_by(ss) %>% dplyr::summarise(first(id), first(Fhrs), first(gear), sum(weight), sum(number))
colnames(st2) <- c("ss", "id", "Fhrs","gear", "weight","number")
st2$CPUEw <- st2$weight/st2$Fhrs
st2$CPUEn <- st2$number/st2$Fhrs

results.table <- data.frame(area=(1:7))
results.table <- cbind(results.table,(subset(st2, gear=="TB") %>% dplyr::group_by(id) %>% dplyr::summarise( mean(CPUEw)))[,2])
results.table <- cbind(results.table,(subset(st2, gear=="GN") %>% dplyr::group_by(id) %>% dplyr::summarise( mean(CPUEw)))[,2])
results.table <- cbind(results.table,(subset(st2, gear=="TB") %>% dplyr::group_by(id) %>% dplyr::summarise( mean(CPUEn)))[,2])
results.table <- cbind(results.table,(subset(st2, gear=="GN") %>% dplyr::group_by(id) %>% dplyr::summarise( mean(CPUEn)))[,2])
results.table <- cbind(results.table,(subset(catch.traits, gear=="TB") %>% dplyr::group_by(id) %>% dplyr::summarise( mean(TrophicLevel, na.rm = TRUE)))[,2])
results.table <- cbind(results.table,(subset(catch.traits, gear=="GN") %>% dplyr::group_by(id) %>% dplyr::summarise( mean(TrophicLevel, na.rm = TRUE)))[,2])
results.table<- cbind(results.table,(subset(catch.traits, gear=="TB") %>% dplyr::group_by(id) %>% dplyr::summarise( mean(as.integer(Gregariousness), na.rm = TRUE)))[,2])
results.table<- cbind(results.table,(subset(catch.traits, gear=="GN") %>% dplyr::group_by(id) %>% dplyr::summarise( mean(as.integer(Gregariousness), na.rm = TRUE)))[,2])
colnames(results.table) <- c("area", "TrapCPUEw", "GillnetCPUEw", "TrapCPUEn", "GillnetCPUEn", "TrophicL_Traps", "TrophicL_Gillnets", "Greg_Traps", "Greg_Gillnets" )

TMean <- subset(catch.traits, gear=="TB" & TrophicLevel!="NA") %>% dplyr::group_by(id, survey_m) %>% dplyr::summarise(TM=mean(na.omit(TrophicLevel)))

#results.table$SpeciesDensity <- sp.area.comb$Average
#results.table$RaoQ <- Raoq$ML_TL[1:7]
results.table$Species_traps <- species.area.table$TB
results.table$Species_gillnets <- species.area.table$GN
results.table$RaoQ_All <- Raoq$All[1:7]
results.table$RaoQ_Traps <- Raoq2$All[1:7]

# formattable(round(results.table, digits = 3))

write.csv2(round(results.table, digits = 3), "./figs/results_table.csv")

results.table %>%
  kable(format = "html", escape = F, digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
  

```

