---
title: "Sudan_catch_analysis"
author: "Erik Olsen"
date: "8/18/2019"
output:
  html_document: default
  pdf_document: default
---
# Analysis of Sudan survey data 2012 - 2013

This R-markdown document contains all r-code used to carry out the analysis for the paper on fish catch rates and species densities (species richness) along the Red Sea coast of Sudan. The analysis is based on three surveys: November 2012, May 2013 and November 2013. 

Here the analysis is structured around the 7 management regions of the Sudanese Red Sea coast as follows:

- Present the region: Bathymetric map
- Sampling locations
- Table of effort (diffent gear types per management areas.)
- Depth distribution of traps by survey and area
- Catch CPUE by area and survey
- Catch CPUE distribution by gear, by species groups (should perhaps do the same for trophic groups)
- CPUE by trophic group pr area and survey
- CPUE by depth and tropich group and survey
- Biodiversity /species densities
- Statistical analysis
  - test of normality shows that we need to use non-parametric and zero-adjusted approaches. 
  - CPUE as a function of area, depth, survey, gear (Zero-adjusted model)

All code and data are stored on GitHub

## Libraries, data etc. 
### Libraries, dependencies and functions
```{r libraries and dependencies, echo=FALSE, warning=FALSE, message=FALSE}
library(reshape2)
library(maps)
library(mapdata)
library(ggplot2)
library(RColorBrewer)
library(sp)
library(rgdal)
library(rgeos)
# plyr mus be loaded befor dplyr for both to work
library(plyr) 
library(dplyr)
library(tidyr)
library(maptools)
library(marmap)
library(classInt)
library(colorspace)
library("formattable")
library("PMCMR")
library(FD)
library("gamlss", lib.loc="/Library/Frameworks/R.framework/Versions/3.5/Resources/library")
```


```{r functions, echo=FALSE}

capwords <- function(s, strict = FALSE) {
  cap <- function(s) paste(toupper(substring(s, 1, 1)),
                           {s <- substring(s, 2); if(strict) tolower(s) else s},
                           sep = "", collapse = " " )
  sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
}
```


### Load and manipulate data
Reading catch, station and traits data. 

Neither station data, nor catch data has a complete depth record, but by combining depth para from each we can get a complete depth parameter for all stations. 
```{r load data, echo=FALSE}
#' Catch data
catch<-read.csv2("catch.csv")
catch$survey<-as.factor(catch$survey)
catch$CPUEw<-catch$weight/catch$Fhrs
catch$CPUEn<-catch$number/catch$Fhrs

#' Station
station <- read.csv2("stations.csv")

#' add depth to catch from station
catch$d1<- station$geardepthstart[match(catch$ss, station$ss)]  
catch$depth <- apply(X=cbind(catch$bdep, catch$d1), MARGIN=1, FUN=max, na.rm=TRUE)

#' Add Month. Year survey column
catch$survey_m<-as.character(catch$survey)
catch$survey_m<-replace(catch$survey_m, grep("2012901", catch$survey_m), "Survey 1: Nov. 2012")
catch$survey_m<-replace(catch$survey_m, grep("2013002", catch$survey_m), "Survey 2: May 2013")
catch$survey_m<-replace(catch$survey_m, grep("2013005", catch$survey_m), "Survey 3: Nov. 2013")

#'Traits
traits <- read.csv2("traits_noNa.csv")
```


### Sudan map and management areas
Loads the map data for Sudan, loads the management areas from shape-file etc. 
```{r map data, echo=FALSE, warning=FALSE}
#' SUDAN MAP DATA
world<- map_data("worldHires", c("Sudan", "Ethiopia")) 
world$survey_m<-c("Survey 1: Nov. 2012")
world2<-world
world2$survey_m<-c("Survey 2: May 2013")
world3<-world
world3$survey_m<-c("Survey 3: Nov. 2013")
world<-rbind(world, world2, world3)

#'  SPATIAL POLYGONS FOR MANAGEMENT AREAS
setwd("/Users/eriko/Documents/GitHub/Sudan2019/Sudan-master")
setwd("./sudan_management_areas")
ManageAreas<-readOGR(dsn = ".", "sudan_regions")
setwd("..")

#' setting correct projection LAT LON
ManageAreas <- spTransform(ManageAreas, CRS("+proj=longlat +ellps=GRS80"))

#' Create  area map with numbers for each region
ManageAreas.f<-fortify(ManageAreas, region="id") #creates X - Y points of the polygons
cnames <- aggregate(cbind(long, lat) ~ id, data=ManageAreas.f, FUN=function(x)mean(range(x)))
cnames$id<-c(1:7)

#' position of Port Sudan
ps<-rbind(data.frame(x=37.21709967,y=19.600512, group=1, survey_m=c("Survey 1: Nov. 2012")),data.frame(x=37.21709967,y=19.600512, group=1, survey_m=c("Survey 2: May 2013") ),data.frame(x=37.21709967,y=19.600512, group=1, survey_m=c("Survey 3: Nov. 2013") ))
```

### Allocating catch positions to management areas
Adds 'Area' to each line in the catch table. 
```{r catch pos to mngmt areas, echo=FALSE}
#' Create spatial.points data base from catch
catch.points<- SpatialPoints(catch[8:9])
proj4string(catch.points) <- proj4string(ManageAreas)
catch.areas<-over(catch.points, ManageAreas)

#' fix area number to area name
alist<-data.frame(name=levels(catch.areas$name), id=c(7,3,2,1,4,5,6))
catch <- cbind(catch, catch.areas[2])
catch$id <- alist$id[match(catch$name, alist$name)]  

#' allocate management areas to catches outside of polygons (choose nearest polygon)
naarea <- catch[is.na(catch$id),]
naarea$id <- naarea$Area
naarea$name[1] <- c("Arakiai")
naarea$name[2:8] <- c("Suakin Archipelago")
naarea$name[9:16] <- c("Donganab")
catch<- rbind(subset(catch, id>=1), naarea) 

#' Change family names to Upper & Lower case
catch$FamGroup2 <- paste(toupper(substr(catch$FamGroup, 1, 1)), tolower(substr(catch$FamGroup, 2, 20)), sep="")
catch$FamGroup2 <- as.factor(catch$FamGroup2)
```

### Modifying dataset and creating subset for stations with catches
```{r nocatch dataset, echo=FALSE}
#' Keep a copy of original 'catch'
catch.unmod <- catch

#' remove Hand Line stations with too short time <1 hr, and too long time >8 hrs
# hist((subset(catch, gear=="HL" & Fhrs>0))$Fhrs, breaks =20) # get an idea of the spread in fishing time
catch <- rbind(subset(catch, gear!="HL"), subset(catch, gear=="HL" & Fhrs>1 & Fhrs<8))

#' selecting only stations with catch & removes stations shorter than 1 hour - two handline stations with fishing time recorded as 5min, that we know is wrong, but we don't have the correct time. 
catch1<-subset(catch, Sci_name!="No catch" & Sci_name!="NO CATCH") 

#' reordering the species groups
fam.names<-c("ACANTHURIDAE", "CARANGIDAE", "CARCHARHINIDAE", "CHIROCENTRIDAE", "LETHRINIDAE", "LUTJANIDAE", "SCOMBRIDAE",     "SERRANIDAE", "OTHER SPP" )  
levels(catch1$FamGroup)<-fam.names

```

### Adding traits to 'catch' data set (without 0-catch stations)
Some more data wrangling that adds the traits from the traits table to the catch data, using only the station with catches (there are no traits for 'NOCATCH' species). 

Also, only select traps, gillnets and handlines as these were the only gear with sufficient numbers and consistent use to be analyzed. 

Lastly adds number of gear deployed at each station to each line. 
```{r add traits,  echo=FALSE}
catch.traits <- catch1
catch.traits$MaxLength<- traits$MaxLength[match(catch1$Sci_name, traits$x1)]  
catch.traits$TrophicLevel<- traits$Trophic.Level[match(catch1$Sci_name, traits$x1)]  
catch.traits$TrophicGroup<- traits$Trophic.group[match(catch1$Sci_name, traits$x1)] 
catch.traits$WaterCol<- traits$Water.column[match(catch1$Sci_name, traits$x1)] 
catch.traits$DielActivity<- traits$Diel.Activity[match(catch1$Sci_name, traits$x1)] 
catch.traits$Habitat<- traits$Habitat[match(catch1$Sci_name, traits$x1)] 
catch.traits$Gregariousness<- traits$Gregariousness[match(catch1$Sci_name, traits$x1)] 
catch.traits$Sci_name_New<- traits$x2[match(catch1$Sci_name, traits$x1)] 
catch.traits$Gregariousness[catch.traits$Sci_name=="Lethrinus mahsena"] <- c(2)
catch.traits$Gregariousness[catch.traits$Sci_name=="Sphyraena jello"] <- c(3)
catch.traits$Gregariousness[catch.traits$Sci_name=="Thunnus albacares"] <- c(3)
catch.traits$Gregariousness[catch.traits$Gregariousness=="#N/A"] <- NA

catch.traits$Gregariousness <- droplevels(catch.traits$Gregariousness)

#' collate Trophic groups and give shorter names
tn<-data.frame(TCat=levels(catch.traits$TrophicGroup), short=c("Invert.", "Herb.", "Coral.", "Carni.", "Plankt.", "Herb."))
catch.traits$TGShort <- tn$short[match(catch.traits$TrophicGroup, tn$TCat)]

#' select only Traps, Gillnet and Handline
catch.traits <- subset(catch.traits, gear=="HL" | gear=="TB" | gear=="GN")
catch.traits$ss1 <- as.factor(catch.traits$ss)

#' add number of stations in that area to each line
c3 <- catch.traits %>% group_by(ss1) %>% summarise( first(depth), first(gear), first(id), first(survey), first(survey_m))
colnames(c3) <- c("ss", "depth", "gear", "id", "survey", "survey_m")
n.area <- c3 %>% group_by(id) %>% summarise(n_distinct(ss))
colnames(n.area) <- c("id", "nos")

catch.traits$NArea <- n.area$nos[match(catch.traits$id, n.area$id)]
```



## Bathymetric map of Sudan
Fig. 1 in MS

A nice, bathymetric map of Sudan with management areas overlaid. 

Currently fonts and colours are adjusted for use on the ICES ASC 2019 poster, removing the bathy-contour lines (se 'lwd' to 0), and varying the colours of management area text to white for areas 1-4. 

```{r bathymap sudan, echo=FALSE}
sudan<-readGEBCO.bathy("./sudan_bathymetry/sudan.nc")
blues <- colorRampPalette(c("midnightblue", "deepskyblue3", "deepskyblue1", "cadetblue3", "cadetblue1", "darkseagreen1", "white"))

tiff(file="./figs/sudan_bathy_map.tiff", width=1900, height=3000, res=400, pointsize=10, compression=c("none"))

plot(sudan, land = TRUE, n = 10, image = TRUE,
     bpal = list(c(min(sudan), -20, "midnightblue", "blue", "lightblue3"),
                 c(-20, 0, "lightblue1", "aquamarine1"),
                 c(0, max(sudan), "gray90", "gray50")), 
     deep = c(-3000, -500, 0),
     shallow = c(-500, -10, 0),
     step = c(500, 200, 0),
     lwd = c(0, 0, 1), lty = c(1, 1, 1),
     col = c("lightgrey", "gray15", "black"),
     drawlabel = c(TRUE, TRUE, TRUE))

scaleBathy(sudan, deg=0.47 ,x="bottomleft", inset=5)

points(37.21709967, 19.600512, pch=19)
points(37.33333, 19.1, pch=19)

text(37.21709967, 19.600512,"Port \nSudan", adj=c(1.2,0), font=2)
text(37.333333, 19.1,"Suakin", adj=c(1.2,0), font=1)
text(37.4, 21.7, "1. Marsas \nnorth of \nDungonab", cex=0.9, font=2, col="white")
text(37.653, 20.92, "2. Dungonab", cex=0.9, font=2, col="white")
text(37.6, 20.3, "3. Arakiai", cex=0.9, font=2, col="white")
text(37.615, 19.95, "4. Port \nSudan", cex=0.9, font=2, col="white")
text(37.6, 18.9, "5. Suakin", cex=0.9, font=2)
text(38, 19.2, "6. Suakin \narchipelago", cex=0.9, font=2)
text(38.3, 18.5, "7. Agig", cex=0.9, font=2)

MApoints<-data.frame(lon=c(37.2, 37.4, 37.38, 37.4, 37.6, 37.9, 38.3), lat=c(21.7, 20.92, 20.3, 19.9, 19, 19.3, 18.5), area=c(1,2,3,4,5,6,7))

ms<-c(1,3,4,5,6,7,8)
for (i in 1:7){
  ma<-subset(ManageAreas.f, id==ms[i])
  polygon(ma$long, ma$lat, lwd=2)  
}

dev.off()
```

![](./figs/sudan_bathy_map.tiff)


## Stations plotted on maps, pr survey
Faceted map plotting position of all catch stations for each of the three surveys. 
```{r station map}
MApoints<-data.frame(lon=c(37.2, 37.53, 37.5, 37.4, 37.3, 37.9, 38.2), lat=c(21.7, 20.92, 20.3, 19.9, 19, 19.3, 18.5), area=c(1,2,3,4,5,6,7))

sudan.map<-ggplot(catch, aes(x=lon, y=lat, group=survey_m)) + geom_point(shape="+", size=6, aes(group=survey_m, colour= survey_m)) + facet_wrap(~survey_m, ncol=3) + geom_path(data=ManageAreas.f,colour="dodgerblue3", aes(x=long, y=lat, group=id)) + geom_polygon(data=world, colour="gray35", fill="gray85", aes(x=long, y=lat, group=survey_m)) +  coord_cartesian(xlim = c(36.5, 39), ylim=c(17.7, 22.5)) + theme_classic() + geom_point(data=ps, size=2, colour="gray35", aes(x=x, y=y)) + geom_text(data=ps, label="PZU", size=3,  hjust=1, vjust=-1.2,  aes(x=x, y=y, group=survey_m)) +geom_text(data=MApoints, aes(x=lon, y=lat, group=area, label=area)) +scale_colour_brewer(type = "seq", palette = "Dark2", name="Survey month & year", labels=c("Nov. 2012", "May. 2013", "Nov. 2013")) + theme(legend.position="none")
sudan.map

ggsave("./figs/station_map.tiff")
```


## Station information
### Station table
Table describing the sampling effort in each area pr survey, number of different gear types, max, min and average depth, number of traps with / without catches. 
```{r station table}
#' TABLE OF SAMPLING EFFORT PR. MANAGEMENT AREA
#' (new for resubmission to PlosOne)
#' --------------------------------------------

s <- unique(catch.unmod$survey)
a <- as.factor(1:7)
ay_info <- data.frame(survey=integer(), id=integer(), Ntraps=integer(), Nhl=integer(), NGn=integer(), TBhrs=double(), HLhrs=double(), GNhrs=double(), DepthAvg=double(), DepthMax=double(), DepthMin=double()) 
cn <- colnames(ay_info)
# depth for TB only (GN are at the surface, and HL not measured)

for (i in 1:length(s)) { 
  y <- subset(catch.unmod, survey==s[i])
  for (j in 1: length(a)){
    ay_1 <- data.frame(survey=integer(), id=integer(), Ntraps=integer(), Nhl=integer(), NGn=integer(), TBhrs=double(), HLhrs=double(), GNhrs=double(), DepthAvg=double(), DepthMax=double(), DepthMin=double())
    ya <- subset(y, id==a[j])
    ay_1 <- rbind(ay_1, c(
      as.integer(as.character(ya$survey[1])), 
      as.integer(as.character(ya$id[1])), 
      length(unique(subset(ya, gear=="TB")$station)),  
      length(unique(subset(ya, gear=="HL")$station)), 
      length(unique(subset(ya, gear=="GN")$station)), 
      sum(subset(ya, gear=="TB")[!duplicated(ya$station),]$Fhrs, na.rm=TRUE),  
      sum(subset(ya, gear=="HL")[!duplicated(ya$station),]$Fhrs, na.rm=TRUE), 
      sum(subset(ya, gear=="GN")[!duplicated(ya$station),]$Fhrs, na.rm=TRUE), 
      mean(subset(ya, gear=="TB")[!duplicated(ya$station),]$depth, na.rm=TRUE), 
      max(subset(ya, gear=="TB")$depth, na.rm=TRUE), 
      min(subset(ya, gear=="TB")$depth, na.rm=TRUE)    ))
    colnames(ay_1) <- cn
    ay_info <- rbind(ay_info, ay_1)
    colnames(ay_info) <- cn
  }
}

formattable(ay_info)

write.csv2(ay_info, "./figs/Area_Station_table.csv")
```


### Number of set traps per depth
Violin plot of depths at wich traps were set. I prefer this visualization of depth ranges of the traps, rather than a (boring) table. 
```{r traps per depth}
c2 <- catch %>% group_by(ss) %>% summarise(first(depth), first(gear), first(id), first(survey), first(survey_m))
colnames(c2) <- c("ss", "depth", "gear", "id", "survey", "survey_m")

dp <- ggplot(subset(c2, gear=="TB" & depth!=0), aes(as.factor(id), depth))

dp <- dp +  geom_violin(scale="width", aes(fill=as.factor(id)), draw_quantiles = c(0.5)) + scale_y_reverse() + labs(x="Management areas", y="Depth (m)") + scale_fill_manual(name="Management area",values=sequential_hcl(n = 7, h = c(0, -100), c = c(80, NA, 40), l = c(53, 75), power = c(1, 1), rev = TRUE, register = ), guide=FALSE ) +theme_bw()  +facet_wrap(~as.factor(survey_m))

dp
tiff(file="./figs/trap_area_depth.tiff", width=3000, height=1500, res=400, pointsize=10, compression=c("none"))
dp
dev.off()


```




## Analysis of catches by traits and other factors

### Catch rates pr area
```{r CPUE area}

#´ Explore: Trap Catch rates (CPUE) by depth and survey
#' not used in MS
#d3 <- ggplot(subset(catch, gear=="TB"), aes(depth, CPUEw))
#d3 + geom_point() + geom_smooth(method="auto")  +facet_wrap(~as.factor(survey_m))

#' main catch by depth plot pr area
ca.plot<-ggplot(subset(catch, gear=="TB" & depth!=0), aes(as.factor(id), CPUEw))

ca.plot <- ca.plot + geom_violin(scale="width", aes(fill=as.factor(id)), draw_quantiles = c(0.5))  + labs(x="Management areas", y="CPUE kg/hrs fishing") + scale_fill_manual(name="Management area",values=sequential_hcl(n = 7, h = c(0, -100), c = c(80, NA, 40), l = c(53, 75), power = c(1, 1), rev = TRUE, register = ), guide=FALSE ) +theme_bw() +facet_wrap(~as.factor(survey_m))
ca.plot

tiff(file="./figs/catch_rates_traps_area_survey.tiff", width=3000, height=2000, res=400, pointsize=10, compression=c("none"))
ca.plot
dev.off()
```

### CPUE by gear type  - by family and trophic group
```{r CPUE gear}
# by genus (families)
catch.3gear<-subset(catch.unmod, gear=="GN" | gear=="HL" | gear =="TB")

gear.plot <- ggplot(catch.3gear, aes(capwords(tolower(as.character(catch.3gear$FamGroup))), fill=survey_m)) + geom_bar(binwidt=0.2, aes( weight=weight)) +theme_bw() + facet_grid(vars(gear), vars(survey_m)) + coord_flip() +xlab("") + ylab("kg")  + guides(fill=guide_legend(title="Survey")) + theme(legend.position = "none")
gear.plot
ggsave("./figs/gearplot_species.tiff")

# by trophic group
gear.plot2 <- ggplot(subset(catch.traits, TGShort!="NA"), aes(TGShort, fill=survey_m)) + geom_bar(binwidt=0.2, aes( weight=weight)) +theme_bw() + facet_grid(vars(gear), vars(survey_m)) + coord_flip() +xlab("") + ylab("kg")  + guides(fill=guide_legend(title="Survey")) + theme(legend.position = "none")
gear.plot2
ggsave("./figs/gearplot_trophic_group.tiff")
```


### CPUE by trophic group by area & by survey (year) 
FIG 5 in Feb 2019 version of MS

Presents the CPUE pr hr, for each management area and survey, standardized by the number of traps set in each area X survey combination. 

Trap catches were dominated by carnivores, occuring in all area X survey combinations, followed by invertivores who also occurred everywhere, albeight in lower densities than the carnivores. 

*Area 2, 5 and 6 seem to have the highest diversity of trophic groups, but this has to be calculated as functional diversity, using the FD package (or similar). *

```{r trophic plot}
trophic.plot <- ggplot(subset(catch.traits, TGShort!="NA" & gear=="TB"), aes(TGShort, CPUEw/NArea))

trophic.plot <- trophic.plot + geom_col(aes(fill=TGShort)) + facet_grid(vars(id), vars(survey_m)) + scale_fill_manual(name="Trophic group",values=qualitative_hcl(n = 5, h = c(0, 295), c = 80, l = 60, register = ) ) + labs(x=" ", y="CPUE (kg/hrs fishing) / (number of traps pr area)") + theme_bw() 
trophic.plot

tiff(file="./figs/trophic_group_cpue.tiff", width=3000, height=3000, res=400, pointsize=10, compression=c("none"))
trophic.plot
dev.off()
```


### Trophic group CPUE by depth
A modified version of Figure 6 in Feb 2019 MS. 

Basically a more complex version of the previous plot. Adds the depth dimension to the plot, and the surveys as colours. 

Shows uniform catch rates of carnivores and invertivores from 0-50m. 

All surveys seem to overlap, so no difference in trophich group depth dependent catch rates between the surveys.

```{r trophic depth plot}
depth.plot <- ggplot(subset(catch.traits, gear=="TB" & TGShort!="NA"), aes(depth, CPUEw)) 

depth.plot <- depth.plot + geom_point(aes(colour=survey_m))  + facet_grid(vars(id),vars(TGShort)) +  scale_colour_manual(name="Survey",values=qualitative_hcl(n = 3, h = c(0, 295), c = 80, l = 60, register = ) ) + labs(x="Depth (m)", y="CPUE (kg(hrs)")  +theme_bw()  

depth.plot

tiff(file="./figs/trophic_group_cpue_depth.tiff", width=3000, height=3000, res=400, pointsize=10, compression=c("none"))
depth.plot
dev.off()
```


### Trophic level in Trap catches
```{r trophic level}
TMean <- subset(catch.traits, gear=="TB" & TrophicLevel!="NA") %>% group_by(id, survey_m) %>% summarise(TM=mean(na.omit(TrophicLevel)))
CMean <- subset(catch.traits, gear=="TB" & TrophicLevel!="NA") %>% group_by(id, survey_m) %>% summarise(CM=mean(CPUEw))
TMean$CM <- CMean$CM

t0.plot <- ggplot(subset(catch.traits, gear=="TB" & TrophicLevel!="NA"), aes(TrophicLevel, CPUEw)) 

t0.plot <- t0.plot + geom_point(aes(colour=TGShort)) + geom_vline(data = TMean, aes(xintercept = TM)) + facet_grid(vars(id),vars(survey_m)) +  scale_colour_manual(name="Trophic group",values=qualitative_hcl(n = 5, h = c(0, 295), c = 80, l = 60, register = ) ) + labs(x="Trophic level", y="CPUE (kg(hrs)")  +theme_bw()  
t0.plot

tiff(file="./figs/trophic_level.tiff", width=3000, height=3000, res=400, pointsize=10, compression=c("none"))
t0.plot
dev.off()
```

### Maxlength in catches
```{r maxlength}
t0.plot <- ggplot(subset(catch.traits, gear=="TB" & TrophicLevel!="NA"), aes(MaxLength, CPUEw)) 

t0.plot <- t0.plot + geom_point(aes(colour=TGShort)) + facet_grid(vars(id),vars(survey_m)) +  scale_colour_manual(name="Trophic group",values=qualitative_hcl(n = 5, h = c(0, 295), c = 80, l = 60, register = ) ) + labs(x="Max lenght(cm)", y="CPUE (kg(hrs)")  +theme_bw()  
t0.plot

tiff(file="./figs/max_length.tiff", width=3000, height=3000, res=400, pointsize=10, compression=c("none"))
t0.plot
dev.off()


```


### Place in water column
```{r place in water column}
catch.traits$WaterCol[catch.traits$WaterCol=="Demersal"] <- "demersal"
tl.plot <- ggplot(subset(catch.traits, WaterCol!="NA" & gear=="TB"), aes(WaterCol, CPUEw/NArea))

tl.plot <- tl.plot + geom_col(aes(fill=WaterCol)) + facet_grid(vars(id), vars(survey_m)) + scale_fill_manual(name="Place in water column",values=qualitative_hcl(n = 6, h = c(0, 295), c = 80, l = 60, register = ) ) + labs(x=" ", y="CPUE (kg/hrs fishing) / number of traps pr area") + theme_bw() + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
tl.plot

tiff(file="./figs/water_col.tiff", width=3000, height=3000, res=400, pointsize=10, compression=c("none"))
tl.plot
dev.off()
```

### Diel activity
```{r diel activity}
t2.plot <- ggplot(subset(catch.traits, DielActivity!="#N/A" & gear=="TB"), aes(DielActivity, CPUEw/NArea))

t2.plot <- t2.plot + geom_col(aes(fill=DielActivity)) + facet_grid(vars(id), vars(survey_m)) + scale_fill_manual(name="Diel activity",values=qualitative_hcl(n = 6, h = c(0, 295), c = 80, l = 60, register = ) ) + labs(x=" ", y="CPUE (kg/hrs fishing) / number of traps pr area") + theme_bw() + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
t2.plot

tiff(file="./figs/diel_act.tiff", width=3000, height=3000, res=400, pointsize=10, compression=c("none"))
t2.plot
dev.off()
```


### Gregariousness
```{r Gregariousness}
t3.plot <- ggplot(subset(catch.traits, Gregariousness!="#N/A" & gear=="TB"), aes(Gregariousness, CPUEw/NArea))

t3.plot <- t3.plot + geom_col(aes(fill=Gregariousness)) + facet_grid(vars(id), vars(survey_m)) + scale_fill_manual(name="Gregariousness",values=qualitative_hcl(n = 6, h = c(0, 295), c = 80, l = 60, register = ) ) + labs(x=" ", y="CPUE (kg/hrs fishing) / number of traps pr area") + theme_bw() + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
t3.plot

tiff(file="./figs/gregariousness.tiff", width=3000, height=3000, res=400, pointsize=10, compression=c("none"))
t3.plot
dev.off()
```




***
## Biodiversity (species densities, trophic diversity etc)


### Aggregated Species density per area for (traps only)
Species per area pr hour of fishing, traps only.

The highest number of species was observed in area 2 (36), followed by area 3, 6 and 7 (25 in each). 

Species density (number of species pr hour of fishing - soak time of traps) ranged from 0.008 (areas 1, 2 and 6) to 0.019 (area 4). Area 4 had the highest species density across all surveys (figure below), while the area with the lowest species density varied by survey: area 1 in November 2012 , area 7 in May 2013 , and area 2 in November 2013 . 
```{r species density}
sp.pr.area <- subset(catch.unmod, gear=="TB") %>% group_by(id, survey_m) %>% summarise(n_distinct(species), n_distinct(ss), sum(Fhrs))
colnames(sp.pr.area) <- c("id", "survey","Nsp","NTraps", "Fhrs")

sp.pr.area$Species_by_NoStations <- (sp.pr.area$Nsp-1) / sp.pr.area$NTraps
sp.pr.area$Species_Hrs <- (sp.pr.area$Nsp-1) / sp.pr.area$Fhrs
write.csv2(sp.pr.area, "./figs/species_density_area.csv")

#calculate species density pr area (all years combined)
sp.pr.area2 <- subset(catch.unmod, gear=="TB") %>% group_by(id) %>% summarise(n_distinct(species), n_distinct(ss), sum(Fhrs))
colnames(sp.pr.area2) <- c("id","Nsp","NTraps", "Fhrs")
sp.pr.area2$Species_by_NoStations <- (sp.pr.area2$Nsp-1) / sp.pr.area2$NTraps
sp.pr.area2$Species_Hrs <- (sp.pr.area2$Nsp-1) / sp.pr.area2$Fhrs
formattable(round(sp.pr.area2, digits=3))

#' Species Density plot
species.plot<-ggplot(sp.pr.area, aes(x=id, y=Species_Hrs)) + geom_bar(stat="identity", fill="steelblue") +theme_bw() + scale_x_reverse(breaks=c(1,2,3,4,5,6,7)) 
sphrs_unmod<- species.plot  + coord_flip() + theme(legend.title = element_text(size=14, face="bold")) + theme(legend.text = element_text(size=12)) + ylab("Number of species caught in traps / median hours of  fishing") + xlab("Management areas from N (1) to S (7)") + facet_wrap(~survey)
sphrs_unmod

ggsave("./figs/species_density.png")
```

### Species density pr station - Zero Adjusted GAM model

The variability and relative difference in rank of species diversity per area as observed in the FIGURE ABOVE was confirmed by the zero-adjusted GAM model (AIC: -4161, df: 12) were all surveys were significant factors in the model. In addition, area 1, 2, 3 and 4 were also significant factors affecting species density, while the non-significant factors were: depth, and areas 5, 6 and 7. 

```{r GAM - species densities}

spue.df <-  subset(catch.unmod, gear=="TB") %>% group_by(ss) %>% summarise(first(survey), first(id), n_distinct(species), first(depth), sum(Fhrs))
colnames(spue.df) <- c("id", "survey", "area", "Nsp", "depth", "Fhrs")
spue.df$Species_Hrs <- spue.df$Nsp / spue.df$Fhrs
#hist(spue.df$spue)

# zero-adjusted model:
sd.mod<-gamlss(Species_Hrs~depth+factor(area)+factor(survey),family=ZAGA, data=spue.df) # AIC: -4161
# gamlss(Species_Hrs~factor(area)+factor(survey),family=ZAGA, data=spue.df)  # AIC: - 4163 - slightly better than the one above, but want to show the one with 'depth' as a factor. 
summary(sd.mod)
```


### Functional diversity
The 'dbFD' function only works when run in-line, not the whole code-chunk. 
 
We used the same approach as Stuart-Smith et al to exclude very few observations of a single species. and excluded all species observations occuring in 2 or fewer stations (traps). 

To avoid dbFD crashing 'm' was set in the 'dbFD' call (the number of PoCA  axes kept as traits to do the FRic analysis). Several levels of 'm' were tested, and this could probably be tested further (to higher levels of 'm')

##### Results for n_spec > 2, m=10
Results based on the output from the dbFD analysis:

No area came out with the highest or lowest FD across all four metrics, but some patterns emerged:

- Area 1 was highest in FEve and FDiv, but lowest in FDis and RaoQ
- Area 2 was consistently high in all metrics, and highest in RaoQ
- Area 3 was high inn all metrics, but never the highest
- Area 4 was lowest in RaoQ, low in FDis, medium in FDiv and FEve
- Area 5 was high in all metrics, but never the highest
- Area 6 was high in FEve, but the lowest in all other three
- Area 7 was lowest in FEve, low in FDis, medium in FDiv and RaoQ

RaoQ is the most used metric (the one used by Stuart-Smith et al), and this is the metric we will use and discuss in the MS. It supports the analysis of species densities for Area 2, although Area 6 has the lowest RaoQ which indicates that although the number of species is higher, their functional diversity is lower (or what we sampled was lower) compared to the other areas. 

Stuart-Smith tests various RaoQs by taking out one trait at a time and this way ranking the traits according to R2.



```{r functional diversity, warning=FALSE}
#ss_list <- levels(catch1$survey)
sp_occur <- catch1  %>% group_by(Sci_name) %>% summarise(n())
colnames(sp_occur) <- c("Sci_name", "n_spec")

# wrangle traits into a traitstable that can be used in FD
missing <- setdiff( x=traits$x1, y=subset(sp_occur, n_spec>2)$Sci_name)
t5 <- traits[!traits$x1 %in% missing,]
tr <- t5[,c(4:10)]
rownames(tr) <- t5$x1
tr[(tr=="#N/A")] <- c(NA)
tr <- droplevels(tr)
tr$diel<-as.integer(tr$Diel.Activity)
tr$diel[(tr$diel==2)] <- c(0)
tr <- tr[,c(1,2,3,4,6,7,8)]

# wrangle 'catch1' to aggregate CPUE pr area (row) pr species (columns)
missing <- setdiff( x=catch1$Sci_name, y=rownames(tr))
c5 <- catch1[!catch1$Sci_name %in% missing,]
c_melt <- melt(c5[,c(28,11,22)], id=c("id", "Sci_name"))

#convert abundance to integer for better calculation performance
med_int1000 <- function(x) as.integer(median(x)*1000)
abund <- dcast(c_melt, id ~ Sci_name, value.var="value", fun=med_int1000) 
# to calculate with raw abundance data, use line below:
#abund <- dcast(c_melt, id ~ Sci_name, value.var="value", fun=median)
abund[is.na(abund)] <- 0
abund <- abund[,c(2:length(abund))]

#calculate FD using dbFD function, m set to 10 to avoid crach (int overload in QHULL)
FuncDiv <- dbFD(tr, abund, corr = c("lingoes"), m=10)

FDt <- as.data.frame(rbind(FuncDiv$FEve, FuncDiv$FDiv, FuncDiv$FDis, FuncDiv$RaoQ))
rownames(FDt) <- c("FEve", "FDiv", "FDis", "RaoQ")
colnames(FDt) <- c("A1", "A2", "A3", "A4", "A5", "A6", "A7")

FDt <- as.data.frame(t(round(FDt, digits=3)))

# create a nice table of FD indicies for the 7 management areas
formattable(FDt, list(    FEve = color_tile("transparent", "lightpink"),     FDiv = color_tile("transparent", "lightpink"),     FDis = color_tile("transparent", "lightpink"),    RaoQ = color_tile("transparent", "lightpink")))

```

#### Rerunning the FD model excluding one trait at a time
Need to reduce 'm' to 9 to avoid crash:

*Zero distance(s)Error in convhulln(tr.FRic, "FA") :* 
  *Received error code 2 from qhull. Qhull error:*
    *QH6114 qhull precision error: initial simplex is not convex. Distance=1.4e-16* 
    
Removing the MaxLenght trait from the model increased R2 by 0.047 compared to the model with all traits included. R2 was reduced when removing any of the other six traits from the model. 

**Conclusion**

- Use the model without MaxLenght
- Removing Maxlength makes sence because the data included a few catches of sharks with very large MaxLengt (>2m) that skews the MaxLenght distribution. Excluding MaxLength creates an analysis around traits that are more common in distribution across species. 
  - Also, the sharks are rowing predators, not so site attached (but that also applies to some pelagic species, especially the large ones, so the same argument applies for these - excluding large body size reduces the importance of these non-site attached species)
```{r FD model excluding traits, warning=FALSE}
# create data frame with original RaoQ and FRic
Raoq <- as.data.frame(c(FuncDiv$RaoQ, FuncDiv$qual.FRic))
colnames(Raoq) <- c("All")
rownames(Raoq) <- c("A1", "A2", "A3", "A4", "A5", "A6", "A7", "R2")


# select tr, but excluding one variable at a time
#tr[,c(1,!2,3,4,5,6,7)]
for (i in 1:7) { 
  FDout <- dbFD(tr[,c(setdiff(1:7, i))], abund, corr = c("lingoes"), m=9)
  raoq1 <- as.data.frame(c(FDout$RaoQ, FDout$qual.FRic))
  colnames(raoq1) <- colnames(tr)[i]
  Raoq <- cbind(Raoq, raoq1)
}
formattable(round(Raoq[8,], digits=3))

formattable(round(Raoq[c(1:7),], digits=3), list(    MaxLength = color_tile("transparent", "lightpink"),     Trophic.Level = color_tile("transparent", "lightpink"),     Trophic.group = color_tile("transparent", "lightpink"),    Water.column = color_tile("transparent", "lightpink"), Habitat = color_tile("transparent", "lightpink"),    Gregariousness = color_tile("transparent", "lightpink"), diel = color_tile("transparent", "lightpink")))

write.csv2(Raoq, "./figs/Raoq_table.csv")

```


***
## Results table
This table combines CPUE, 
```{r results table}
st2 <- catch %>% group_by(ss) %>% summarise(first(id), first(Fhrs), first(gear), sum(weight))
colnames(st2) <- c("ss", "id", "Fhrs","gear", "weight")
st2$CPUEw <- st2$weight/st2$Fhrs

results.table <- data.frame(area=(1:7))
results.table <- cbind(results.table,(subset(st2, gear=="TB") %>% group_by(id) %>% summarise( mean(CPUEw)))[,2])
results.table <- cbind(results.table,(subset(st2, gear=="GN") %>% group_by(id) %>% summarise( mean(CPUEw)))[,2])
results.table <- cbind(results.table,(subset(catch.traits, gear=="TB") %>% group_by(id) %>% summarise( mean(TrophicLevel, na.rm = TRUE)))[,2])
results.table <- cbind(results.table,(subset(catch.traits, gear=="GN") %>% group_by(id) %>% summarise( mean(TrophicLevel, na.rm = TRUE)))[,2])
results.table<- cbind(results.table,(subset(catch.traits, gear=="TB") %>% group_by(id) %>% summarise( mean(as.integer(Gregariousness), na.rm = TRUE)))[,2])
results.table<- cbind(results.table,(subset(catch.traits, gear=="GN") %>% group_by(id) %>% summarise( mean(as.integer(Gregariousness), na.rm = TRUE)))[,2])
colnames(results.table) <- c("area", "TrapCPUE", "GillnetCPUE", "TrophicL_Traps", "TrophicL_Gillnets", "Greg_Traps", "Greg_Gillnets" )

TMean <- subset(catch.traits, gear=="TB" & TrophicLevel!="NA") %>% group_by(id, survey_m) %>% summarise(TM=mean(na.omit(TrophicLevel)))

results.table$SpeciesDensity <- sp.pr.area2$Species_Hrs
results.table$RaoQ <- Raoq$MaxLength[1:7]

formattable(round(results.table, digits = 3))

write.csv2(round(results.table, digits = 3), "./figs/results_table.csv")

```


***
## Statistical analysis CPUE catches
### Testing for normality of CPUE catches
Shapiro test for normality and Q-Q plots shows that data is non-normal and zero-inflated. 
```{r normality test}
catch.df<-subset(catch1, number>0 & weight>0)

#' test for normality
shapiro.test(catch.df$CPUEw)
shapiro.test(catch.df$CPUEn)

#' Q-Q plots
library("qqplotr")
gg <- ggplot(data = catch.df, mapping = aes(sample = CPUEw,  fill = factor(id))) + stat_qq_band() + stat_qq_line() + stat_qq_point(size=0.2) + facet_wrap(~ id, scales="free") +  labs(x = "CPUEw predicted", y = "CPUEw Observed") + ggtitle("Q-Q plots CPUEw")
gg
```
### Test for difference in CPUE between areas (non-parametric test)
Both CPUEw and CPUEn are significantly different between the areas, but only CPUEn in the pairwise tests (nemenyi post-hoc) for areas 4:3 and 6:4. THis points to that Area 4 is the 'odd one out' in terms of CPUE along the Sudanses coast. 

Significant difference for CPUEW between May 2013 survey and the Nov 2012 and Nov 2013 surveys, but not between the Nov12 and Nov13 surveys

```{r nonpara CPUA, eval=FALSE}
#### NOT RUN BECAUSE ZERO-ADJUSTED GAM MODEL IS BETTER!

#' 1 - test for difference between areas
kruskal.test(catch.df$CPUEw, catch.df$id)
kruskal.test(catch.df$CPUEn, catch.df$id)

#' Post-Hoc tests of Kruskal - Wallis results
library("PMCMR", lib.loc="/Library/Frameworks/R.framework/Versions/3.5/Resources/library")
posthoc.kruskal.nemenyi.test(catch.df$CPUEw, catch.df$id, "Chisq")
posthoc.kruskal.nemenyi.test(catch.df$CPUEn, catch.df$id, "Chisq")
#' significant difference in CPUEn between area 6 and area 4

#' 2 test for differences between surveys
kruskal.test(catch.df$CPUEw, catch.df$survey)
kruskal.test(catch.df$CPUEn, catch.df$survey)

posthoc.kruskal.nemenyi.test(catch.df$CPUEw, catch.df$survey, "Chisq")
posthoc.kruskal.nemenyi.test(catch.df$CPUEn, catch.df$survey, "Chisq")
```

### Zero-inflated model (better approach)
This method is more appropriate than the Kruskal-Wallis approach above as it evaluates the effects of all factors at the same time, instead of evaluating the effects of area separate from survey. 

Developed a model for CPUEw including depth, area, survey and gear, which performed better than a simpler model excluding gear. 

Depth is not significant

Significant factors:

* Area (1, 5, 6 and 7)
* Surveys (all)
* Gear (all gear types)
```{r zero-infl model}
#' Fitting GAM Zero Inflated model to catch data

# select only variables used in model
c2 <- catch[,c(1,14,22,25,28)]
c2 <- subset(c2, gear=="TB" | gear=="GN" |gear=="HL")

# CPUEw - GAMLSS ZAGA model incluing survey, area, depth & gear 
mod<-gamlss(CPUEw~depth+factor(id)+factor(survey)+factor(gear),family=ZAGA, data=c2) 
summary(mod)
```







